---
title: "Split-Treat Data Preparation"
author: "Dan Aurell"
date: "2023-10-27"
output:   
  html_document: 
    keep_md: yes
---


# Workflow 

## What data do I have from the app?

App data - d77
-
Shaw 5/24/2022
Wire 5/31/2022
Shooting Range 6/7/2022
Turfgrass 6/11/2022

Post-harvest - no app data
-
Shaw 7/1/2022
Wire 7/2/2022
Shooting Range 7/2/2022
Turfgrass 7/2/2022




# Setup

```{r}
library(tidyverse)
library(lme4)
library(stringr)
```


# Read in data

## Read csv
```{r}
paper <- read.csv("./data/manual_merge_paper_2023-05-23 v2.csv")
colony <- read.csv("./data/2022-06-14_125633__techteam__all_sample_export.csv", skip = 9, na.strings = c("", " ", "NA"))
frame <- read.csv("./data/2022-06-14_125650__techteam__frame_side_export.csv")


```

## Pare down and rename columns

Paper data
```{r}
paper <- paper %>% 
  rename(colony_num = split_ids)
```

Frame-level data
```{r}

frame2 <- frame %>% 
  rename(
    project = frame__box__sample__sample_event__project,  
    colony_num = frame__box__sample__colony_num,
    date = frame__box__sample__sample_event__collection_date,
    box_num = frame__box__box_number,
    box_size = frame__box__size,
    frame_num = frame__frame_number,
    side = side_name,
    bees = bee_coverage,
    cap_wk = capped_worker_brood_coverage,
    cap_dr = capped_drone_brood_coverage,
    honey = honey_coverage,
    pollen = pollen_coverage,
    extra_bees = frame__box__bee_coverage,
    inspection_order = frame__box__sample__inspection_order
         )


frame2 <- frame2 %>% 
  select(
        project,
    colony_num,
    date,
    box_num,
    box_size,
    frame_num,
    side,
    bees,
    cap_wk,
    cap_dr,
    honey,
    pollen,
    extra_bees,
    inspection_order,
         id,
         frame,
         notes
         )

frame2 <- frame2 %>% 
  filter(date %in% c("2022-05-24", "2022-05-31", "2022-06-07", "2022-06-11")) 


# Quality control
unique(frame2$colony_num)

 frame2 %>% 
  group_by(date, colony_num) %>% 
  summarize(n = n()) %>% 
   arrange(desc(n))

```
## Deal with "extra bees" so that it is counted only once per box
```{r}
frame2 %>% 
  group_by(date, colony_num) %>% 
  summarize(n = n()) # There are 36-72 rows per colony and date


frame2$extra_bees_temp <- frame2$extra_bees # clone column
frame2$extra_bees <- NA # Erase the values in original column

frame2 <- frame2 %>% 
  mutate(extra_bees = ifelse(frame_num == 1 & side == "A", extra_bees_temp, NA)) %>% 
  select(-extra_bees_temp)

# Replace NA with 0
frame2$bees[is.na(frame2$bees)] <- 0
frame2$extra_bees[is.na(frame2$extra_bees)] <- 0

# Replace 2520 with 25
frame2$bees[frame2$bees == 2520] <- 0

frame2 <- frame2 %>% 
  mutate(bees_total = bees + extra_bees)

```


Correct resource coverages for different box/frame sizes - express all as percentage of one side of a Deep frame
```{r}
# Convert to Deep-frame equivalents
frame2 <- frame2 %>% mutate(
  conversion = ifelse(box_size == "Medium", 0.64, 1)
)

names(frame2)

# append "_C" to a column to mean converted
frame2 <- frame2 %>% 
  mutate(
    bees_C = bees*conversion,
    cap_wk_C = cap_wk*conversion,
    cap_dr_C = cap_dr*conversion,
    honey_C = honey*conversion,
    pollen_C = pollen*conversion
  )
```

## Summarize into a "resources" data frame that sums resources for each colony, since the colony is the experimental unit

Also convert from percentages of frame-sides to frames of [resource]

```{r}
resources <- frame2 %>% 
  group_by(colony_num) %>% 
  # 100 would mean that there is one complete side of bees
  
  # Divide by 100 to get "sides"
  # Then divide by 2 to get "frames"

  # In total, divide by 200
  
  summarise(
    frames_bees = sum(bees_C, na.rm = TRUE)/200,
    frames_wbr = sum(cap_wk_C, na.rm = TRUE)/200,
    frames_dbr = sum(cap_dr_C, na.rm = TRUE)/200,
    frames_honey = sum(honey_C, na.rm = TRUE)/200
    )
```


Colony-level data
```{r}

names(colony)

colony2 <- colony %>% 
  select(sample_event__project,
         sample_event__collection_date,
         colony_num,
         deeps_num,
         mediums_num,
         threequarters_num,
         shallows_num,
         supers_num,
         queen_status,
         frames_of_bees,
         efb_observed
         ) %>% 
  rename(project = sample_event__project,
         collection_date = sample_event__collection_date
         )
```



# Understand data and how to analyze it

How to identify just relevant app data for day 77
"colony" and "frame" are the colony and frame level app data
```{r}
unique(colony$project)

pattern <- "date"
  grep(pattern, names(colony), value=TRUE)  

pattern <- "coverage"
  grep(pattern, names(frame), value=TRUE)  
  
unique(colony$collection_date)
```

# Prep data 

## Prep colony level app data
```{r}
# Colony2 is just selected and renamed

colony2 %>% 
  filter(collection_date %in% c("2022-05-24", "2022-05-31", "2022-06-07", "2022-06-11")) %>% 
  group_by(collection_date, colony_num) %>% 
  summarize(n())

# colony3 is filtered to the data of interest
colony3 <- colony2 %>% 
  filter(collection_date %in% c("2022-05-24", "2022-05-31", "2022-06-07", "2022-06-11")) %>% 
  filter(!is.na(queen_status)) %>% # removes some duplicates introduced by app
  arrange(colony_num)

unique(colony3$colony_num) #138 including the blank space
# But a few duplicated colonies exist

# Replace 21-043 with 22-043
colony3$colony_num[colony3$colony_num == "21-034"] <- "22-034"

# focus on queen status and efb columns
colony4 <- colony3 %>% 
  select(colony_num, queen_status, efb_observed)

# Rename queen status and efb columns
colony4 <- colony4 %>% 
  rename(ch77_qstatus = queen_status,
         ch77_efb = efb_observed
         )


# Remove duplicate rows introduced by the app
colony4 <- colony4[!duplicated(colony4[, 1]), ]

# "pa" means "paper_app"
pa <- left_join(paper, colony4)

# Make more pleasant treatment names in a trt_label column
pa <- pa %>% 
  mutate(
    trt_label = if_else(trt == "1-Control", "Control",
              if_else(trt == "2-Apivar", "Apivar",
                    if_else(trt == "4-Dribble18", "OA Dribble", 
                            if_else(trt == "6-ManyDribble", "5x OA Dribble",
                                   if_else(trt == "5-Vapor18", "OA Vapor",
                                           if_else(trt == "7-HopGuard", "HopGuard",
                                                   if_else(trt == "3-Taktic", "Amitraz E.C.", "NA"
                                                           )
                                                   )))))))
# Add a trt_type column
pa <- pa %>% 
  mutate(
    trt_type = if_else(trt == "1-Control", "Control",
              if_else(trt == "2-Apivar", "Synthetic",
                    if_else(trt == "4-Dribble18", "Natural", 
                            if_else(trt == "6-ManyDribble", "Natural",
                                   if_else(trt == "5-Vapor18", "Natural",
                                           if_else(trt == "7-HopGuard", "Natural",
                                                   if_else(trt == "3-Taktic", "Synthetic", "NA"
                                                           )
                                                   )))))))

pa <- pa %>%
     mutate(trt_label = fct_relevel(trt_label, "Control", "Apivar", "Amitraz E.C.", "OA Dribble", "5x OA Dribble", "OA Vapor", "HopGuard"))

```

## Prep pa data

Tag colonies that had EFB by day 77
```{r}

pa$efb_columns <- paste(
  pa$reason_removal, 
  pa$ch77_efb # This did not add any additional ones
  )

pa$efb <- ifelse(grepl("EFB", pa$efb_columns), 1, 0)



# Number of EFB colonies per treatment
pa %>%
  filter(efb == 1) %>% 
  group_by(trt) %>% 
  summarize(n = n()) # Very even across groups
```

Impute bees in alcohol wash for 3 colonies on day 77
```{r}
pa %>%filter(ch77_wash_workers == "Missing")
# 22-078
# 22-108
# 22-124

missed <- c("22-078", "22-108", "22-124")

# Set as numeric... these were "character" because of "Missing" entry for bees in sample columns
pa$ch77_wash_workers <- as.numeric(pa$ch77_wash_workers)
pa$ch77_wash_drones <- as.numeric(pa$ch77_wash_drones)
pa$ch77_wash_queens <- as.numeric(pa$ch77_wash_queens)

  
pa <- pa %>%   
  mutate(ch2_wash_bees = ch2_wash_workers + ch2_wash_drones + ch2_wash_queens,
         ch77_wash_bees = ch77_wash_workers + ch77_wash_drones + ch77_wash_queens,
         chpost_wash_bees = chpost_wash_workers + chpost_wash_drones + chpost_wash_queens)

# Here doing some descriptive stats (that will be reported in paper) - early because I need it to impute bee number in sample

pa_piv <- pa %>% 
  pivot_longer(
    c(ch2_wash_bees, ch77_wash_bees, chpost_wash_bees), 
    names_to = "day", 
    values_to = "bees_in_sample") %>% 
  select(colony_num, day, bees_in_sample)


pa_piv %>% 
  group_by(day) %>% 
  summarize(
    mean = mean(bees_in_sample, na.rm=TRUE),
    min = min(bees_in_sample, na.rm=TRUE),
    max = max(bees_in_sample, na.rm=TRUE)
  )

pa %>%
  mutate(
    
  )


# To impute day-77 
pa <- pa %>% 
  mutate(
    ch77_wash_bees = ifelse(colony_num %in% missed, 362, ch77_wash_bees)
    )

```





Determine queen issues including day 77
```{r}
# Survival to day 77 - assess based on absence of other issues

# Number of queen events per treatment up to and including day 72

# failed to mate
  # look in column reason_removal
# drone layer
# queenless
  # If QL or DL present in columns ch44_qstatus, ch58_qstatus, ch72_qstatus, 


pa$qstatuses <- paste(
  pa$ch44_qstatus, 
  pa$ch58_qstatus, 
  pa$ch72_qstatus, 
  pa$ch77_qstatus,
  pa$reason_removal
  )

pa$failed <- ifelse(grepl("Failed", pa$reason_removal), 1, 0)
pa$weak <- ifelse(grepl("Weak|Small", pa$reason_removal), 1, 0)
pa$QL <- ifelse(grepl("QL|VS|VQS|VQ", pa$qstatuses), 1, 0)
pa$DL <- ifelse(grepl("DL", pa$qstatuses), 1, 0)

# This initial categorization is not entirely correct, since some queens that were laying eggs on day 30 were not actually laying worker eggs on that day. Instead of DL they should be tagged "failed" 

```


# Refine categorization of queen data - Queen establishment success Y/N

Criterion: did they begin laying worker eggs by Day 30?
- If not by that time we culled them

Data to use: all
- For other variables that relate to colony development over time, we exclude colonies that ever developed EFB
- Since this is a short-term measure, we can include these EFB colonies
- First time EFB was observed was d44
- What time frame to look at...

Dates to look at
- Day 30 data is the most key... 
- Simplistically we could just look and see were they QR on day 30 or not?
- But a few colonies had young brood on day 30, so we can't necessarily tell yet if they are drone layers - so let's also look at day 44 data
- It is only ambiguous if the brood age was L3 or lower on day 30
  - L3 on day 30 corresponds to anything with a layday of 24 or higher


```{r results = "hide"}
# Exploring how to assess 

pa %>% 
  filter(start_layday <= 23)
  # Of the 138 colonies that had a layday of 23 or lower, all were QR/QS on d30 and none were DL on d44

# All colonies whose queens started laying on day 23 or before, did successfully lay worker eggs


pa %>% 
  filter(start_layday >= 24) 
  # Of the 27 colonies that had a layday of 24 or higher...
pa %>% 
  filter(start_layday >= 24) %>% 
  filter(ch44_qstatus == "DL")
  # ...8 were DL on d44
```

Refined criterion for "successfully established":
- Laying worker brood on day 30 (only able to asses for layday <=23)
- If start_layday >=24, then consider established if QR/QS at ch44_qstatus

```{r}
# Success column means "did they ever lay worker eggs"
pa$established <- "NA"

# Nested ifelse to implement criterion
pa <- pa %>% 
  mutate(
    established = ifelse(ch30_qstatus %in% c("QR", "QS") & start_layday <= 23, 1, 
                     ifelse(start_layday >= 24 & ch44_qstatus %in% c("QR", "QS"), 1, 0))
  )

# Correct the failed column
pa$failed <- ifelse(pa$established == 0, 1, 0)

# When we a queen never laid worker eggs, this "failed" needs to override the other issues (overrides 8 DL and one QL)
pa$QL <- ifelse(pa$failed == 1, 0, pa$QL)
pa$DL <- ifelse(pa$failed == 1, 0, pa$DL)

pa <- pa %>% 
  mutate(issues = failed + QL + DL + efb + weak)

# 60 colonies had any issue

# I have also confirmed that each colony only has one issue.


```


## Add resources data to pa data
```{r}
pa <- left_join(pa, resources)
```

Correct the hive weight based on the number of supers and 12 lbs per super
12.005 was mean (min-max: 11.1-12.8) of 20 supers weighed before putting them on

```{r}
12.005*0.454 # = 5.45
11.1*0.454 # 5.04
12.8*0.454 # 5.8

pa <- pa %>% 
  mutate(post_lbs_C = post_lbs - 12*(post_supers_use-1),
         post_kgs_C = 0.454*post_lbs_C)
```



# Write an_dat for use in future steps
```{r}
an_dat <- pa

write.csv(an_dat, "../2_Analysis/data_prepared/an_dat.csv")
```


