---
title: "Split-Treat Descriptive Stats"
author: "Dan Aurell"
date: "2023-10-27"
output:   
  html_document: 
    keep_md: yes
---

```{r}
library(MASS)
library(tidyverse)
# library(stats)
library(lme4)
```

```{r}
an_dat <- read.csv("./data_prepared/an_dat.csv")
```


# Descriptive stats - including some plotting

## Overall establishment rate and total colonies removed due to each issue
```{r}
sum(an_dat$failed) #23
sum(an_dat$established) #154

# 154/177 = 87.0 %

sum(an_dat$QL) # 9
sum(an_dat$DL) # 4
sum(an_dat$efb) # 19
sum(an_dat$weak) # 5

# 23+9+4+19+5 = 60

sum(an_dat$issues) # 60

# Most constrained dataset includes 117 colonies of the initial 177


```
## Establishment rate by treatment
```{r}
established_summary <- an_dat %>% 
  group_by(trt_label) %>% 
  summarize(n = n(),
            established_num = sum(established),
            established_p = established_num/n
          )

established_summary <- established_summary %>% 
  mutate(failed_num = n - established_num,
    failed_p = 1-established_p)
  
```

### Analyze establishment rate by treatment - using a GLM or GLMM framework

```{r}

## GLM framework

m1 <- glm(established ~ trt_label, family = binomial, data = an_dat)
m0 <- glm(established ~ 1, family = binomial, data = an_dat)

summary(m1) 
summary(aov(m1)) # To generate an overall p value
# No significant difference in establishment rate across groups (F 6, 170 = 1.24, P=0.288)

## GLMM framework

m2 <- glmer(established ~ trt_label + (1|new_yard), family = binomial, data=an_dat)
# singular
m2.0 <- glmer(established ~ 1 + (1|new_yard), family = binomial, data=an_dat)

summary(m2)

anova(m2, m2.0)
# No significant difference in establishment rate across groups (ChiSq 6 = 10.365, P=0.1101)

```

### Analyze issue rate by treatment - using a GLM or GLMM framework

```{r}

## GLM framework

m3 <- glm(issues ~ trt_label, family = binomial, data = an_dat)

summary(aov(m3)) # To generate an overall p value
# No significant difference in occurrence of issues between treatments (F 6, 170 = 0.794, P=0.576)


## GLMM framework

m4 <- glmer(issues ~ trt_label + (1|new_yard), family = binomial, data=an_dat)
# singular
m4.0 <- glmer(issues ~ 1 + (1|new_yard), family = binomial, data=an_dat)

anova(m4, m4.0)
# No significant difference in occurrence of issues between treatments (ChiSq 6 = 5.275, p=0.5091) 

```



### Analyze establishment rate by treatment - using a GLMM framework

```{r}
```




### Plot establishment rate by treatment
```{r}

established_summary_piv2 <- established_summary %>% 
  pivot_longer(c(established_p, failed_p), names_to = "outcome", values_to = "proportion") 

established_summary_piv2 <- established_summary_piv2 %>% 
  mutate(
    outcome_label = if_else(outcome == "established_p", "Established",
      if_else(outcome == "failed_p", "Did not establish", NA))
  )

# Relevel factors for plotting
established_summary_piv2 <- established_summary_piv2 %>%
     mutate(trt_label = fct_relevel(trt_label, "Control", "Apivar", "Amitraz E.C.", "OA Dribble", "5x OA Dribble", "OA Vapor", "HopGuard")
                        )

established_summary_piv2 <- established_summary_piv2 %>%
     mutate(outcome_label = fct_relevel(outcome_label, "Did not establish", "Established")
            )


# Plot establish vs. not
established_summary_piv2 %>% 
  ggplot(aes(x = trt_label, y = proportion, fill = outcome_label)) +
  geom_col(position = "stack", color = "black") +
  scale_colour_manual(
  limits = c("Did not establish", "Established"),
  labels = c("Did not establish", "Established"),
  values = c("white", "black"), 
  aesthetics = "fill"
) +
  scale_y_continuous(limits = c(0,1)) + 
  ylab("Proportion of colonies") +
  theme_classic(base_size = 30) + 
  theme(axis.text.x=element_text(angle=45,hjust=1), axis.title.x=element_blank(), legend.title = element_blank())
  
# ggsave("./outputs/established_p_2024-03-02 v2.png", width = 12, height = 8.625, units = "in")
```




## Colony issues

```{r}
# Issues

issues_summary <- an_dat %>% 
  group_by(trt_label) %>% 
  summarize(n = n(),
            
            efb_num = sum(efb),
            efb_p = efb_num/n,
            
            weak_num = sum(weak),
            weak_p = weak_num/n,
            
            failed_num = sum(failed),
            failed_p = failed_num/n,

            QL_num = sum(QL),
            QL_p = QL_num/n,
            
            DL_num = sum(DL),
            DL_p = DL_num/n
            
          )



```


```{r}
sum(an_dat$issues) #60 colonies had any issue
# i.e. 117 had no issue
# 117/177 i.e., 66.1 % of colonies had no issue

# Check if this lines up with the number of colonies for which I have Varroa data on day 77. It does. 

```

### Prep data for colony issues graphing

```{r}
issues_summary_pivoted <- issues_summary %>% 
  pivot_longer(c(efb_p, weak_p, failed_p, QL_p, DL_p), names_to = "reason", values_to = "proportion") 


issues_summary_pivoted <- issues_summary_pivoted %>% 
  mutate(
    reason_label = if_else(reason == "efb_p", "EFB",
      if_else(reason == "weak_p", "Weak",
                           if_else(reason == "failed_p", "Did not establish",
                                   if_else(reason == "DL_p", "Drone layer",
                                           if_else(reason == "QL_p", "Queenless", "NA"
                           )
    
    
    )))))

# Relevel factors for plotting

issues_summary_pivoted <- issues_summary_pivoted %>%
     mutate(trt_label = fct_relevel(trt_label, "Control", "Apivar", "Amitraz E.C.", "OA Dribble", "5x OA Dribble", "OA Vapor", "HopGuard"),
            reason_label = fct_relevel(reason_label,
    "Weak",
    "EFB",
    "Queenless",
    "Drone layer", 
    "Did not establish"
    )
            )





```

### Plotting colony issues

```{r}
# Graph with queen issues, "weak" and "EFB"

issues_summary_pivoted %>% 
  ggplot(aes(x = trt_label, y = proportion, fill = reason_label)) +
  geom_col(position = "stack") +
  scale_fill_viridis_d(direction = -1) +
  scale_y_continuous(limits = c(0,1)) + 
  ylab("Proportion of colonies") +
  theme_classic(base_size = 30) + 
  theme(axis.text.x=element_text(angle=45,hjust=1), axis.title.x=element_blank(), legend.title = element_blank())
  
ggsave("./outputs/CoD 2024-03-02 v4.png", width = 12, height = 8.625, units = "in")


# Graoh Only efb, weakness

issues_summary_pivoted %>% 
  filter(reason %in% c("weak_p", "efb_p")) %>% 
  ggplot(aes(x = trt_label, y = proportion, fill = reason_label)) +
  geom_col(position = "stack") +
  scale_fill_viridis_d(direction = -1) +
  scale_y_continuous(limits = c(0,1)) + 
  ylab("Proportion of colonies") +
  theme_classic(base_size = 30) + 
  theme(axis.text.x=element_text(angle=45,hjust=1), axis.title.x=element_blank(), legend.title = element_blank())
  
ggsave("./outputs/CoD 2024-03-02 v5.png", width = 12, height = 8.625, units = "in")

# Only queen issues

issues_summary_pivoted %>% 
  filter(reason %in% c("failed_p", "QL_p", "DL_p")) %>% 
  ggplot(aes(x = trt_label, y = proportion, fill = reason_label)) +
  geom_col(position = "stack") +
  scale_fill_viridis_d(direction = -1) +
  scale_y_continuous(limits = c(0,1)) + 
  ylab("Proportion of colonies") +
  theme_classic(base_size = 30) + 
  theme(axis.text.x=element_text(angle=45,hjust=1), axis.title.x=element_blank(), legend.title = element_blank())
  
ggsave("./outputs/CoD 2024-03-02 v6.png", width = 12, height = 8.625, units = "in")

```


# Deeper queens analysis - date they started laying

There's no need to statistically analyze this. There's no biological basis to hypothesize that treatments would increase the establishment success of queens.

There is biological basis to hypothesize that treatments might harm queens, thus decreasing establishment success.

No treatment had lower success than Control, so we just state that there was no indication that any treatment lowered establishment success.


# Determine length of brood break for each colony

```{r}
# It's only sensible to look at the length of brood break for colonies which DID successfully establish. Thus filter to success = "Y"

# A worker brood cell becomes receptive to Varroa on day 8, and emerges on day 21

an_dat <- an_dat %>% 
  # filter(established == 1) %>% 
  # select(colony_num, trt_label, start_layday) %>% 
  mutate(age_start_lay = start_layday - 2,
          emerge_day = 21, # Age when worker brood starts emerging
         receptive_day = start_layday + 8,
         days_break = receptive_day - emerge_day,
         completeYN = ifelse(days_break > 0, "Complete", "Not Complete"),
         completeYN = fct_relevel(completeYN, "Not Complete", "Complete")
         )

# Relevel factors for plotting
an_dat <- an_dat %>%
     mutate(trt_label = fct_relevel(trt_label, "Control", "Apivar", "Amitraz E.C.", "OA Dribble", "5x OA Dribble", "OA Vapor", "HopGuard")
                        )


# Describe overall results - for colonies whose queens established

table(an_dat$completeYN[an_dat$established==1])
# Of 154 colonies whose queens successfully established, 101 experienced a complete brood break (65.6%), whereas 53 did not experience a complete brood break

 
an_dat %>% 
  filter(established == 1) %>% 
  summarise(
    mean = mean(days_break, na.rm = T),
    min = min(days_break, na.rm = T),
    max = max(days_break, na.rm = T)
  )


# Considering worker brood only, the overall mean length of brood break was 3.5 days, with some colonies experiencing an overlap of as much as 4 days of old queen's brood and new queen, and some colonies experiencing as much as 15 days of absence of receptive brood.


```


```{r}
# Some basic plots to visualize data
an_dat %>% 
  ggplot(aes(x=days_break)) +
  geom_histogram(binwidth = 1)

an_dat %>% 
  ggplot(aes(x=days_break)) +
  geom_density()

an_dat %>%
  ggplot(aes(y=trt_label, x=days_break)) +
  geom_violin()


# Working on a final figure

## Graph option 1 - preferred
an_dat %>%
  ggplot(aes(x=trt_label, y=days_break)) + # Length of brood break
  geom_boxplot(outlier.shape = NA) +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed", alpha = 0.5) +
  geom_jitter(alpha=0.2, width = 0.2, height = 0.1) +
  stat_summary(fun.y=mean, geom="point", shape=20, size=5, alpha = 0.5, color = "red") +
  scale_x_discrete(limits = rev) +
  ylab("Length of brood break (days)") +
  theme_classic(base_size = 18) +
  theme(axis.title.y=element_blank()) +
  coord_flip()


ggsave("./outputs/length_break_2024-03-03 v1.png", width = 12, height = 8.625, units = "in")
```

```{r}
## Graph option 2 - not preferred
an_dat %>%
  ggplot(aes(x=trt_label, y=days_break)) + # Length of brood break
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(color = completeYN), alpha=0.4, width = 0.2, height = 0.1) +
  stat_summary(fun.y=mean, geom="point", shape=20, size=5, alpha = 0.5) +
  scale_x_discrete(limits = rev) +
  ylab("Length of brood break (days)") +
  theme_classic(base_size = 18) +
  theme(axis.title.y=element_blank(), legend.position = "none") +
  coord_flip()
```

To potentially look at:
1. Statistically analyze do any treatments cause delay in queen establishment?
  - Note that brood break length has what may be a bimodal distribution - which would make sense if some of the late colonies' queens derive from an emergency cell, not the one we placed. Will cause difficulties to use brood break length or establishment time as a response variable.
1a) Are some queens our cell queens and others are emergency queens?  
  - One way to pose question: What proportion of queens in each treatment established before 18 days post-placement?
2. Use brood break length as a predictor of efficacy against Varroa


# Finish by writing analytical dataset back to data_prepared folder... so that queen data is now included


```{r}
write.csv(an_dat, "./data_prepared/an_dat_q.csv")
```




