---
title: "Analysis - queens, survival"
author: "Dan Aurell"
date: "2023-05-28"
output:   
  html_document: 
    keep_md: yes
---

# THE USEFUL PARTS OF THIS ANALYSIS HAVE ALREADY BEEN MIGRATED TO:
1.1 Data Preparation
2.1 Descriptive Stats

# Setup

```{r}
library(tidyverse)
library(lme4)
library(car)
```

# Read in data

```{r}
an_dat <- read.csv("./data_prepared/an_dat.csv")

an_dat <- an_dat %>%
  mutate(trt_label = fct_relevel(trt_label, "Control", "Apivar", "Amitraz E.C.", "OA Dribble",  "5x OA Dribble","OA Vapor", "HopGuard"))

```

# Queen establishment success Y/N

Criterion: did they begin laying worker eggs by Day 30?
- If not by that time we culled them

Data to use: all
- For other variables that relate to colony development over time, we exclude colonies that ever developed EFB
- Since this is a short-term measure, we can include these EFB colonies
- First time EFB was observed was d44
- What time frame to look at...

Dates to look at
- Day 30 data is the most key... 
- Simplistically we could just look and see were they QR on day 30 or not?
- But a few colonies had young brood on day 30, so we can't necessarily tell yet if they are drone layers - so let's also look at day 44 data
- It is only ambiguous if the brood age was L3 or lower on day 30
  - L3 on day 30 corresponds to anything with a layday of 24 or higher


```{r results = "hide"}
# Exploring how to assess 

an_dat %>% 
  filter(start_layday <= 23)
  # Of the 138 colonies that had a layday of 23 or lower, all were QR/QS on d30 and none were DL on d44

# All colonies whose queens started laying on day 23 or before, did successfully lay worker eggs


an_dat %>% 
  filter(start_layday >= 24) 
  # Of the 27 colonies that had a layday of 24 or higher...
an_dat %>% 
  filter(start_layday >= 24) %>% 
  filter(ch44_qstatus == "DL")
  # ...8 were DL on d44
```

Refined criterion:
- Laying worker brood on day 30 (only able to asses for layday <=23)
- If start_layday >=24, then consider established if QR/QS at ch44_qstatus

## Determine which queens successfully established
```{r}
# names(an_dat)

# Make a manageable dataset
establish_dat <- an_dat %>% 
  select(colony_num,
         new_yard,
         trt_label,
         ch18_eggs,
         ch18_qstatus,
         ch18_ageoldest,
         ch30_eggs,
         ch30_qstatus,
         ch30_ageoldest,
         ch44_eggs,
         ch44_qstatus,
         start_layday
         )

# Success column means "did they ever lay worker eggs"
establish_dat$success <- "NA"

# Nested ifelse to implement criterion
establish_dat <- establish_dat %>% 
  mutate(
    success = ifelse(ch30_qstatus %in% c("QR", "QS") & start_layday <= 23, "Y", 
                     ifelse(start_layday >= 24 & ch44_qstatus %in% c("QR", "QS"), "Y", "N"))
  )

```


## Summarize what proportion of queens successfully established
- Per treatment
- Overall

```{r}
# Overall
establish_dat %>% 
  group_by(success) %>% 
  summarise(n = n()) %>% 
  spread(success, n)
# Overall, 23 failed and 154/177 successfully established

success_all_tab <- data.frame(trt_label = "All", Y = 154, N=23)


# By treatment
success_trt_tab <- establish_dat %>% 
  group_by(trt_label, success) %>% 
  summarise(n = n()) %>% 
  spread(success, n) %>% 
  ungroup()

success_trt_tab <- success_trt_tab %>% 
  mutate(N = ifelse(is.na(N), 0, N)) %>% 
  select(trt_label, Y, N)

success_tab <- rbind(success_all_tab, success_trt_tab)

# Set the order it will print in
success_tab <- success_tab %>%
  mutate(trt_label = fct_relevel(trt_label, "All", "Control", "Apivar", "Amitraz E.C.", "OA Dribble",  "5x OA Dribble","OA Vapor", "HopGuard"))

success_tab <- success_tab %>% 
  mutate(total = Y + N,
         perc = 100*Y/total
         )

success_tab %>% 
  arrange(trt_label)

```









