---
title: "2_2X Analysis Varroa Exploratory"
author: "Dan Aurell"
date: "2024-02-01"
output:   
  html_document: 
    keep_md: yes
---


# Determine how length of brood break might influence efficacy *within treatment*...

The difficulty is that we can't calculate efficacy on an every-hive basis since it will divide by zero for some hives.

So then we use an alternate method

d77 Varroa ~ d2 Varroa + brood_break_length

I would want to analyze d2 Varroa as a predictor in a standard way, but to analyze the predictor brood break length in a more flexible way since I do not assume a linear relationship.

I like how ggplot can do a loess regression, but how to do apply the nonlinear regression only to one variable?

Generalized additive models?


Prepare data for GAM
```{r}
temp <- temp %>% 
  mutate(age_start_lay = start_layday - 2,
          emerge_day = 21,
         receptive_day = start_layday + 8,
         days_break = receptive_day - emerge_day,
         completeYN = ifelse(days_break > 0, "Complete", "Not Complete"),
         completeYN = fct_relevel(completeYN, "Not Complete", "Complete")
         )
```


```{r}
 pattern <- "var"
   grep(pattern, names(temp), value=TRUE)  

m6 <- gam(ch77_var_total ~ s(days_break) + ch2_var_total, data = temp, subset = trt)
m7 <- gam(ch77_var_total ~ s(days_break) + ch2_var_total, data = temp)

m8 <- gam(ch77_var_total ~ s(days_break, by = trt_label) + ch2_var_total, data = temp)

summary(m8)
```


