---
title: "Analysis - Varroa"
author: "Dan Aurell"
date: "2024-01-07"
output:   
  html_document: 
    keep_md: yes
---

# Setup

```{r}
library(scales)
library(MASS)
library(tidyverse)
library(emmeans)
library(broom)
library(multcomp)
library(lme4)
library(lubridate)
library(mgcv)
library(ggpubr) # For combo plot
library(gt)
```

# Read in data

```{r}
an_dat <- read.csv("./data_prepared/an_dat.csv")

an_dat %>% 
  filter(issues == 0)


names(an_dat)
```

# Prepare data for Varroa infestation analysis

Exclude any colonies that had EFB or queen issues or were removed due to weakness by day 77
```{r}
temp <- an_dat %>% 
  filter(issues == 0) %>% 
  mutate(ch2_wash_bees = ch2_wash_workers + ch2_wash_drones + ch2_wash_queens,
         ch77_wash_bees = ch77_wash_workers + ch77_wash_drones + ch77_wash_queens,
         chpost_wash_bees = chpost_wash_workers + chpost_wash_drones + chpost_wash_queens,
         ch2_perc_var = 100*ch2_var_total/ch2_wash_bees,
         ch77_perc_var = 100*ch77_var_total/ch77_wash_bees,
         chpost_perc_var = 100*chpost_var_total/chpost_wash_bees
         )

# 117 of 177 colonies still in trial on d77: 66% of colonies. Not too bad.

names(temp)

unique(temp$trt_label)

temp <- temp %>%
  mutate(trt_label = fct_relevel(trt_label, "Control", "Apivar", "Amitraz EC", "OA Dribble",  "5x OA Dribble","OA Vapor", "HopGuard"))

# calculate the number of days between day 2 sampling and post sampling
temp <- temp %>% 
  mutate(
    num_days = mdy(chpost_date) - mdy(ch2_date),
    days_numeric = as.numeric(num_days)
  )

unique(temp$days_numeric) # 95 to 113 days
mean(unique(temp$days_numeric)) #102.75 ; round to 100

# pattern <- "date"
#   grep(pattern, names(temp), value=TRUE)  

temp_post <- 
  temp %>% 
  filter(is.na(post_issues))

```

# Assess were there differences in Varroa load between treatments?

## Describe arithmetic means on day 2, 77, post
```{r}
temp %>% 
  summarise(mean = mean(ch2_perc_var))

temp %>% 
  summarise(mean = mean(ch77_perc_var))

temp %>% 
  summarise(mean = mean(chpost_perc_var))

temp %>%
  group_by(trt_label) %>% 
    summarise(mean2 = mean(ch2_perc_var),
              mean77 = mean(ch77_perc_var),
              meanpost = mean(chpost_perc_var)
              )


# Greatest numerical difference between 5x OA dribble and HopGuard

# 2.0474221/0.8906983 = 2.3 times as many mites in the HopGuard group as in the 5x OA dribble. This justifies correcting for it statistically.
```


## Make long version of Varroa data
```{r}
temp_piv <- temp %>% 
  pivot_longer(c(ch2_perc_var, ch77_perc_var, chpost_perc_var), names_to = "day", values_to = "perc_var")

temp_piv <- temp_piv %>% mutate(post_issues = ifelse(is.na(post_issues), 0, post_issues))

temp_piv <- temp_piv %>% 
  filter(
    !((post_issues == 1) & (day == "chpost_perc_var"))
  )
# excluded post data for 10 colonies that had issues 


temp_piv <- temp_piv %>%
  mutate(trt_label = fct_relevel(trt_label, "Control", "Apivar", "Amitraz EC", "OA Dribble",  "5x OA Dribble","OA Vapor", "HopGuard"))

# Rename "day" values to 2, 77, post

temp_piv <- temp_piv %>% 
  mutate(day = if_else(day == "ch2_perc_var", "2", 
                       if_else(day == "ch77_perc_var", "77", 
                               if_else(day == "chpost_perc_var", "Post", NA))))

```

## Determine how many were above/below 3% threshold for each treatment each day

```{r}
temp_piv <- temp_piv %>%
  mutate(above = if_else(perc_var >= 3, 1, 0))

# This above/below information is now saved in a df
above_df <- temp_piv %>% 
  group_by(day, trt_label) %>% 
  summarize(n=n(),
            above_num = sum(above),
            above_perc = 100*above_num/n,)

```

## Graph Varroa levels on day 2, 77, post

Graph it
```{r}
temp_piv %>% 
  ggplot(aes(x=trt_label, y = perc_var, fill = trt_label)) +
  geom_boxplot() +
  ylab("Varroa infestation of\nadult bees (%)") +
  facet_wrap(~day) +
  theme_classic(base_size = 30) +
  theme(axis.text.x=element_text(angle=45,hjust=1), axis.title.x=element_blank(), legend.position = "none")

# ggsave("./outputs/supp_varroa_daily 2024-04-24.png", width = 12, height = 8.625, units = "in")
```

## Assess pre-treatment differences statistically

```{r}

## Including only non-issues colonies

m.day2 <- glm.nb(ch2_var_total~trt_label, data = temp)
summary(m.day2)

m.day2.0 <- glm.nb(ch2_var_total~1, data = temp)
summary(m.day2.0)

anova(m.day2, m.day2.0) # Pre-treatment differences were not significant (ChiSq 6 = 4.023; P=0.674)

## Qualitatively same result if include all colonies
m.day2 <- glm.nb(ch2_var_total~trt_label, data = an_dat)
summary(m.day2)

m.day2.0 <- glm.nb(ch2_var_total~1, data = an_dat)
summary(m.day2.0)

anova(m.day2, m.day2.0) # Pre-treatment differences were not significant (P=0.84)


```
# Post-treatment differences

```{r}
an_dat %>% 
  ggplot(aes(x=trt_label, y = ch77_var_total)) +
  geom_boxplot()
# Looks like t

m.day77 <- glm.nb(ch77_var_total~trt_label, data = an_dat)
summary(m.day2)

m.day77.0 <- glm.nb(ch77_var_total~1, data = an_dat)
summary(m.day2.0)

anova(m.day77, m.day77.0) # Day 77 differences were significant (ChiSq 6 = 28.9; P=6.30e-05)
```



# Run models for d77 varroa

## I would consider including yard as a random effect since we blocked by yard
This should increase power to detect treatment effects

But many give singular fits, and the estimates are so similar to when you omit random effect, that I will not use random effects





## Correcting for number of bees in sample by including it as an offset

```{r}
# Omitting random effect
m3 <- glm.nb(ch77_var_total ~ 
                 trt_label + 
                 ch2_perc_var +
                 offset(log(ch77_wash_bees)), 
               data=temp)
# summary(m3)
summary(m3)

m0 <- glm.nb(ch77_var_total ~ 
                 # trt_label + 
                 ch2_perc_var + 
                 offset(log(ch77_wash_bees)), 
               data=temp)
summary(m0)

anova(m3, m0) 
# Treatment significantly influenced Varroa load on day 77 (ChiSq 6 = 37.27869, P=1.553757e-06=0.000001553757)



summary(m3)# Very similar estimates of effect

```



## How did the Varroa infestations vary between treatments on day 77?
Used to make statistical comparisons between treatment groups

```{r}
# Make a priori contrasts between treatments
K <- contrMat(table(temp$trt_label), type="Tukey") # all contrasts
Ksub <- rbind(K[c(3,4,5,6,8,9,10,11,16,17),]) # A priori hypotheses
Ksub2 <- rbind(K[c(1:6),]) # Curious what happens if I just compare each treatment to control

summary(glht(m3, linfct = mcp(trt_label = Ksub)),
        test = adjusted("single-step"))

# Steury comment: Approach varies journal to journal
# If I do it this way that I originally planned (a priori contrasts), I will need to justify why I'm selecting some comparisons in particular...

# One thing I might do is just run all post-hoc comparisons
  # If qualitatively the same results as my initial approach I could just use those

summary(glht(m3, linfct = mcp(trt_label = K)),
        test = adjusted("single-step"))
# Yes, this does provide qualitatively the same results. 
# I use this for reporting results

# Curious how much influence the p-value correction has
summary(glht(m3, linfct = mcp(trt_label = K)),
        test = adjusted("none"))

# I want to be conservative and make sure I'm not unnecessarily writing off OA vapor. If I test each treatment against control, is OA still stat similar to control?

summary(glht(m3, linfct = mcp(trt_label = Ksub2)),
        test = adjusted("single-step"))
# Yes, OA is still not distinguishable from control.



```

Several natural treatments had lower Varroa levels than the control:
HopGuard was significantly lower than control (z=-4.150, p< 0.001), as was 5x OA Dribble (z=-4.435, p< 0.001) and OA Dribble (z=-3.518, p0.00787). However, the difference between Control and OA Vapor was not significant (z=-1.494, p=0.74787).

The only other significant difference between treatments was that OA Vapor had more Varroa mites than Amitraz EC (z=3.443, P=0.01030).

A - Control
AB - OA Vapor
BC - OA dribble
BC- 5x OA dribble
BC - HopGuard
BC - Apivar
C - AEC


## Final code used for efficacy comparisons
```{r}
# Used for contrasts in LMS study
# mo2021W <- glht(m2021W, linfct = mcp(OperationType="Tukey"))
# t2021W <- tidy(mo2021W)


cont77adj <- summary(glht(m3, linfct = mcp(trt_label = K)),
        test = adjusted("single-step"))

cont77raw <- summary(glht(m3, linfct = mcp(trt_label = K)),
        test = adjusted("none"))

cont77adj_t <- tidy(cont77adj)
cont77raw_t <- tidy(cont77raw)

cont77all <- full_join(cont77adj_t, cont77raw_t)

cont77all$adj.p.value <- sprintf("%.3f", cont77all$adj.p.value)
cont77all$p.value <- sprintf("%.3f", cont77all$p.value)
cont77all$statistic <- sprintf("%.2f", cont77all$statistic)
# After the fact will manually change 0.000 to < 0.001

```

## Make the table of contrasts
```{r}
cont77all2 <- cont77all %>% 
  select(contrast, statistic, adj.p.value, p.value)


cont_tbl <- gt(cont77all2)

cont_rtf <- cont_tbl %>%
  as_rtf()

my_conn <- file("./outputs/SuppTable3_2024-04-24.RTF", "w")
writeLines(cont_rtf, my_conn)
close(my_conn)


```





# Run models for post-honey Varroa
The 4 experimental blocks had a staggered start, but ended at the same time (due to the honey harvest )
To account for the different length of time that each yard was in the experiment, I include the number of days as a predictor, and make predictions normalized to ~103 days of experiment length.




```{r}

m4 <- glm.nb(chpost_var_total ~ 
                 trt_label + 
                 ch2_perc_var + 
               days_numeric +
                 offset(log(chpost_wash_bees)), 
               data=temp_post)

summary(m4)


m4.0 <- glm.nb(chpost_var_total ~ 
                 trt_label + 
                 ch2_perc_var + 
               # days_numeric +
                 offset(log(chpost_wash_bees)), 
               data=temp_post)

summary(m4.0)

anova(m4, m4.0)
# Effect of days was significant p = 0.02; go with m4

m4.00 <- glm.nb(chpost_var_total ~ 
                 # trt_label + 
                 ch2_perc_var + 
               days_numeric +
                 offset(log(chpost_wash_bees)), 
               data=temp_post)

summary(m4.00)
anova(m4, m4.00)
# Effect of treatment was significant on day 103 as well (ChiSq 6 = 26.6724; P=0.0001667661) 


```

## How did the Varroa infestations vary between treatments on day 103?
Used to make statistical comparisons between treatment groups

```{r}
# Make a priori contrasts between treatments
K <- contrMat(table(temp$trt_label), type="Tukey")
Ksub <- rbind(K[c(3,4,5,6,8,9,10,11,16,17),])

summary(glht(m4, linfct = mcp(trt_label = Ksub)),
        test = adjusted("single-step"))

summary(glht(m4, linfct = mcp(trt_label = K)),
        test = adjusted("single-step"))

```


# Summarize results as efficacy
- And get confidence intervals around efficacy

## Day 77 Efficacy - no random effect of yard included

```{r}
tidy(m3)
summary(m3)

eff77_3 <- tidy(m3)

eff77_3CI <- as.data.frame(confint(m3))
eff77_3b <- cbind(eff77_3, eff77_3CI)
eff77_3b <- eff77_3b %>% 
  rename(LCL = "2.5 %", UCL = "97.5 %")



eff77_3b <- eff77_3b %>% 
  mutate(times_as = exp(estimate),
         efficacy = 100*(1-times_as),
         
         times_as_UCL = exp(UCL),
         efficacy_LCL = 100*(1-times_as_UCL),
         
         times_as_LCL = exp(LCL),
         efficacy_UCL = 100*(1-times_as_LCL)
         )

eff77_3b <- eff77_3b[2:7,]


# Make more pleasant treatment names in a trt_label column
eff77_3b <- eff77_3b %>% 
  mutate(
    trt_label = if_else(term == "trt_labelApivar", "Apivar",
                    if_else(term == "trt_labelOA Dribble", "OA Dribble", 
                            if_else(term == "trt_label5x OA Dribble", "5x OA Dribble",
                                   if_else(term == "trt_labelOA Vapor", "OA Vapor",
                                           if_else(term == "trt_labelHopGuard", "HopGuard",
                                                   if_else(term == "trt_labelAmitraz EC", "Amitraz EC", "NA"
                                                           )
                                                   ))))))
# Add a trt_type column
eff77_3b <- eff77_3b %>% 
  mutate(
    trt_type = if_else(trt_label == "Apivar", "Synthetic",
                    if_else(trt_label == "OA Dribble", "Natural", 
                            if_else(trt_label == "5x OA Dribble", "Natural",
                                   if_else(trt_label == "OA Vapor", "Natural",
                                           if_else(trt_label == "HopGuard", "Natural",
                                                   if_else(trt_label == "Amitraz EC", "Synthetic", "NA"
                                                           )
                                                   ))))))

eff77_3b <- eff77_3b %>% 
  mutate(
    trt_label = fct_relevel(trt_label, "Apivar", "Amitraz EC", "OA Dribble", "5x OA Dribble", "OA Vapor", "HopGuard")
  )

eff77_3b$day <- 77

```



Mixed model eff77_2b has near-identical estimates of efficacy and CIs around efficacy as the standard model.
I will choose to use the non-mixed efficacy


## Day 103 Efficacy

```{r}
library(broom)

# Using the non-mixed model

effpost <- tidy(m4)
effpost_CI <- as.data.frame(confint(m4))

effpost <- cbind(effpost, effpost_CI)
effpost <- effpost %>% 
  rename(LCL = "2.5 %", UCL = "97.5 %")


effpost <- effpost %>% 
  mutate(times_as = exp(estimate),
         efficacy = 100*(1-times_as),
         
         times_as_UCL = exp(UCL),
         efficacy_LCL = 100*(1-times_as_UCL),
         
         times_as_LCL = exp(LCL),
         efficacy_UCL = 100*(1-times_as_LCL)
         )

effpost <- effpost[2:7,]

## To Do - clean up this data frame like I did for the eff77 data frame.

# Make more pleasant treatment names in a trt_label column
effpost <- effpost %>% 
  mutate(
    trt_label = if_else(term == "trt_labelApivar", "Apivar",
                    if_else(term == "trt_labelOA Dribble", "OA Dribble", 
                            if_else(term == "trt_label5x OA Dribble", "5x OA Dribble",
                                   if_else(term == "trt_labelOA Vapor", "OA Vapor",
                                           if_else(term == "trt_labelHopGuard", "HopGuard",
                                                   if_else(term == "trt_labelAmitraz EC", "Amitraz EC", "NA"
                                                           )
                                                   ))))))
# Add a trt_type column
effpost <- effpost %>% 
  mutate(
    trt_type = if_else(trt_label == "Apivar", "Synthetic",
                    if_else(trt_label == "OA Dribble", "Natural", 
                            if_else(trt_label == "5x OA Dribble", "Natural",
                                   if_else(trt_label == "OA Vapor", "Natural",
                                           if_else(trt_label == "HopGuard", "Natural",
                                                   if_else(trt_label == "Amitraz EC", "Synthetic", "NA"
                                                           )
                                                   ))))))

effpost <- effpost %>% 
  mutate(
    trt_label = fct_relevel(trt_label, "Apivar", "Amitraz EC", "OA Dribble", "5x OA Dribble", "OA Vapor", "HopGuard")
  )


effpost$day <- 103
```


Get predictions for day 77
```{r}
pred_77 <- emmip(m3, ~ trt_label, type = "response", CIs = TRUE, plotit = FALSE, offset = log(100))
# this is based on the non-included variable (Varroa infestation on day 2) being set to its mean... 
# Predictions are now set to the value of offset = 100... (mites per 100 bees)

pred_77$day <- 77

# Check this against the actual means per group
temp %>% 
  group_by(trt_label) %>% 
  summarise(varroa = mean(ch77_perc_var))
# Matches closely



```

Get predictions for day 103

```{r}
pred_post <- emmip(m4, ~ trt_label, type = "response", CIs = TRUE, plotit = FALSE, offset = log(100), at = list(ch2_perc_var = 1.42))

pred_post$day <- 103

```


```{r}
# Combine day 2, 77, 103 into one plottable data frame

mean(temp$ch2_perc_var) # 1.42 Varroa per 100 bees

pred_pre <- data.frame(pred_77$trt_label, yvar = c(rep(1.42, 7)), day = c(rep(2, 7)))
pred_pre <- pred_pre %>% 
  rename(trt_label = pred_77.trt_label)

  

  

pred <- bind_rows(pred_pre, pred_77, pred_post)
```


```{r}
# Check sample size per treatment and day
ssize77 <- temp %>% 
  group_by(trt_label) %>% 
  summarize(n=n())

ssize2 <- ssize77

ssize103 <- temp_post %>% 
  group_by(trt_label) %>% 
  summarize(n=n())

ssize2$day <- 2
ssize77$day <- 77
ssize103$day <- 103

ssize <- rbind(ssize2, ssize77, ssize103)


```


# Make table of Varroa infestation predictions

```{r}
# Combine predictions and sample size

pred2 <- left_join(pred, ssize)

pred2 <- pred2 %>% 
  select(day, trt_label, n, yvar, LCL, UCL)

# Day 103 predictions are now based on a day 2 level of 1.42 - by adding at argument in emmip

pred2$yvar <- sprintf("%.2f", pred2$yvar)
pred2$LCL <- sprintf("%.2f", pred2$LCL)
pred2$UCL <- sprintf("%.2f", pred2$UCL)


pred2 <- pred2 %>% 
  mutate(
    Estimate = paste0(yvar, " [", LCL, "-", UCL, "]")
  ) %>% 
  select(-c(yvar, LCL, UCL))

pred3<- pred2 %>% 
  group_by(day) %>% 
  arrange(trt_label, .by_group = TRUE) %>% 
  ungroup()

# Note that
```


```{r}
pred_tbl <- gt(pred3)

pred_rtf <- pred_tbl %>%
  as_rtf()

my_conn <- file("./outputs/SuppTable1_2024-04-24.RTF", "w")
writeLines(pred_rtf, my_conn)
close(my_conn)
```


# Make table of efficacy predictions

```{r}
# Combine day 77 and day 103 efficacy

eff_all <- rbind(eff77_3b, effpost)


eff_all2 <- eff_all %>% 
  select(day, trt_label, efficacy, efficacy_LCL, efficacy_UCL)

eff_all2$efficacy <- sprintf("%.2f", eff_all2$efficacy)
eff_all2$efficacy_LCL <- sprintf("%.2f", eff_all2$efficacy_LCL)
eff_all2$efficacy_UCL <- sprintf("%.2f", eff_all2$efficacy_UCL)


eff_all2 <- eff_all2 %>% 
  mutate(
    Efficacy = paste0(efficacy, " [", efficacy_LCL, "-", efficacy_UCL, "]")
  ) %>% 
  select(-c(efficacy, efficacy_LCL, efficacy_UCL))

eff_all3<- eff_all2 %>% 
  group_by(day) %>% 
  arrange(trt_label, .by_group = TRUE) %>% 
  ungroup()

```

```{r}
eff_all_tbl <- gt(eff_all3)

eff_all_rtf <- eff_all_tbl %>%
  as_rtf()

my_conn <- file("./outputs/SuppTable2_2024-04-24.RTF", "w")
writeLines(eff_all_rtf, my_conn)
close(my_conn)
```

# Graph Varroa infestation predictions

```{r}
#extract hex color codes for a plot with three elements in ggplot2 
hex <- hue_pal()(6)

#overlay hex color codes on actual colors
show_col(hue_pal(h = c (25,290), h.start = -15)(6))
hex <- hue_pal(h = c (25,290), h.start = -15)(6)
hex
```

DAN work on colors for color and BW version of combo figure

```{r}

# User-selected color scale

paneA <- pred %>% 
  ggplot(mapping = aes(y = yvar, x = day, colour = trt_label)) +
  geom_hline(yintercept = 3, linetype = 2, colour = "red") +
  geom_line(size = 1.5, alpha = 0.8) +
  geom_point(size = 3, alpha = 0.8) +

  scale_colour_manual(
  limits = c("Control", "OA Vapor", "OA Dribble", "Apivar", "5x OA Dribble", "HopGuard","Amitraz EC"),
  labels = c("Control", "OA Vapor", "OA Dribble", "Apivar", "5x OA Dribble", "HopGuard","Amitraz EC"),
  values = c("black", "#619CFF", "#619CFF", "#F8766D", "#619CFF", "#619CFF", "#F8766D"), 
  aesthetics = "color") +
  
  scale_x_continuous(breaks = c(2, 77, 103)) +
  scale_y_continuous(breaks = c(0,1,2,3,4,5,6,7), labels = c(0,"","",3,"","",6,""), limits = c(0,7)) +
  
  ylab("Varroa infestation of\nadult bees (%)") +
  xlab("Experiment day") +
  theme_classic(base_size = 30)
  
  


paneA.lab <- 
  paneA +
  theme(legend.title=element_blank())
  

# ggsave("./outputs/varroa_pred_2024-03-11 v1-manual.png", width = 12, height = 8.625, units = "in")


paneA.nolab <- 
  paneA +
  theme(legend.position = "none")


# BW color scale

# Make paneA.BW

paneA.BW <- pred %>% 
  ggplot(mapping = aes(y = yvar, x = day, colour = trt_label)) +
  geom_hline(yintercept = 3, linetype = 2) +
  geom_line(size = 1.5, alpha = 0.8) +
  geom_point(size = 3, alpha = 0.8) +

  scale_colour_manual(
  limits = c("Control", "OA Vapor", "OA Dribble", "Apivar", "5x OA Dribble", "HopGuard","Amitraz EC"),
  labels = c("Control", "OA Vapor", "OA Dribble", "Apivar", "5x OA Dribble", "HopGuard","Amitraz EC"),
  values = c("black", "grey50", "grey50", "grey75", "grey50", "grey50", "grey75"), 
  aesthetics = "color") +
  
  scale_x_continuous(breaks = c(2, 77, 103)) +
  scale_y_continuous(breaks = c(0,1,2,3,4,5,6,7), labels = c(0,"","",3,"","",6,""), limits = c(0,7)) +
  
  ylab("Varroa infestation of\nadult bees (%)") +
  xlab("Experiment day") + 
  theme_classic(base_size = 30)



paneA.BW.lab <- 
  paneA.BW +
  theme(legend.title=element_blank())


paneA.BW.nolab <- 
  paneA.BW +
  theme(legend.position = "none")


```


## Plot d77 efficacy

```{r}
# Color, no label
paneB <- eff77_3b %>% 
  ggplot(aes(y = efficacy, x = trt_label, fill = trt_type)) + 
  geom_bar(stat = "identity", color = "black") +
  geom_errorbar(aes(ymin=efficacy_LCL, ymax=efficacy_UCL), width=0.5) +
  scale_colour_manual(
    limits = c("Control", "Synthetic", "Natural"),
    labels = c("Control", "Synthetic", "Natural"),
    values = c("black", "#F8766D", "#619CFF"), 
    aesthetics = "fill"
  ) +  
  ylab("Efficacy (%)") +
  scale_y_continuous(breaks = c(-20, 0,20,40,60,80,100)) +
  theme_classic(base_size = 30) +
  
  theme(axis.text.x=element_text(angle=45,hjust=1),  axis.title.x=element_blank(), legend.position = "none")

# color, label_1

eff77_3b %>% 
  ggplot(aes(y = efficacy, x = trt_label, fill = trt_type)) + 
  geom_bar(stat = "identity", color = "black") +
  geom_errorbar(aes(ymin=efficacy_LCL, ymax=efficacy_UCL), width=0.5) +
  scale_colour_manual(
    limits = c("Control", "Synthetic", "Natural"),
    labels = c("Control", "Synthetic", "Natural"),
    values = c("black", "#F8766D", "#619CFF"), 
    aesthetics = "fill"
  ) +  
  ylab("Efficacy (%)") +
  scale_y_continuous(breaks = c(-20, 0,20,40,60,80,100)) +
  theme_classic(base_size = 30) +
  
  theme(axis.text.x=element_text(angle=45,hjust=1),  axis.title.x=element_blank(), legend.title = element_blank())

ggsave("./outputs/efficacy_C2 2024-12-21 label_1.png", width = 12, height = 8.625, units = "in")

# color, label_2

eff77_3b %>% 
  ggplot(aes(y = efficacy, x = trt_label, fill = trt_type)) + 
  geom_bar(stat = "identity", color = "black") +
  geom_errorbar(aes(ymin=efficacy_LCL, ymax=efficacy_UCL), width=0.5) +
  scale_colour_manual(
    limits = c("Synthetic", "Natural"),
    labels = c("Synthetic", "Natural"),
    values = c("#F8766D", "#619CFF"), 
    aesthetics = "fill"
  ) +  
  ylab("Efficacy (%)") +
  scale_y_continuous(breaks = c(-20, 0,20,40,60,80,100)) +
  theme_classic(base_size = 30) +
  
  theme(axis.text.x=element_text(angle=45,hjust=1),  axis.title.x=element_blank(), legend.title = element_blank())

ggsave("./outputs/efficacy_C2 2024-12-21 label_2.png", width = 12, height = 8.625, units = "in")

# BW, no label

paneB.BW <- eff77_3b %>% 
  ggplot(aes(y = efficacy, x = trt_label, fill = trt_type)) + 
  geom_bar(stat = "identity", color = "black") +
  geom_errorbar(aes(ymin=efficacy_LCL, ymax=efficacy_UCL), width=0.5) +
  scale_colour_manual(
    limits = c("Control", "Synthetic", "Natural"),
    labels = c("Control", "Synthetic", "Natural"),
    values = c("black", "grey75", "grey50"), 
    aesthetics = "fill"
  ) +  
  ylab("Efficacy (%)") +
  scale_y_continuous(breaks = c(-20, 0,20,40,60,80,100)) +
  theme_classic(base_size = 30) +
  
  theme(axis.text.x=element_text(angle=45,hjust=1),  axis.title.x=element_blank(), legend.position = "none")




# BW, label_1

eff77_3b %>% 
  ggplot(aes(y = efficacy, x = trt_label, fill = trt_type)) + 
  geom_bar(stat = "identity", color = "black") +
  geom_errorbar(aes(ymin=efficacy_LCL, ymax=efficacy_UCL), width=0.5) +
  scale_colour_manual(
    limits = c("Control", "Synthetic", "Natural"),
    labels = c("Control", "Synthetic", "Natural"),
    values = c("black", "grey75", "grey50"), 
    aesthetics = "fill"
  ) +  
  ylab("Efficacy (%)") +
  scale_y_continuous(breaks = c(-20, 0,20,40,60,80,100)) +
  theme_classic(base_size = 30) +
  
  theme(axis.text.x=element_text(angle=45,hjust=1),
        axis.title.x=element_blank(), legend.title = element_blank())

ggsave("./outputs/efficacy_BW 2024-12-21 label_1.png", width = 12, height = 8.625, units = "in")

# BW, label_2

eff77_3b %>% 
  ggplot(aes(y = efficacy, x = trt_label, fill = trt_type)) + 
  geom_bar(stat = "identity", color = "black") +
  geom_errorbar(aes(ymin=efficacy_LCL, ymax=efficacy_UCL), width=0.5) +
  scale_colour_manual(
    limits = c("Synthetic", "Natural"),
    labels = c("Synthetic", "Natural"),
    values = c("grey75", "grey50"), 
    aesthetics = "fill"
  ) +  
  ylab("Efficacy (%)") +
  scale_y_continuous(breaks = c(-20, 0,20,40,60,80,100)) +
  theme_classic(base_size = 30) +
  
  theme(axis.text.x=element_text(angle=45,hjust=1),
        axis.title.x=element_blank(), legend.title = element_blank())

ggsave("./outputs/efficacy_BW 2024-12-21 label_2.png", width = 12, height = 8.625, units = "in")
  
```

## Plot d103 efficacy

New plot
```{r}
effpost %>% 
  ggplot(aes(y = efficacy, x = trt_label, fill = trt_type), alpha = 0.6) + 
  geom_bar(stat = "identity", color = "black") +
  geom_errorbar(aes(ymin=efficacy_LCL, ymax=efficacy_UCL), width=0.5) +
  scale_colour_manual(
    limits = c("Control", "Synthetic", "Natural"),
    labels = c("Control", "Synthetic", "Natural"),
    values = c("white", "#F8766D", "#619CFF"), 
    aesthetics = "fill"
  ) +  
  ylab("Efficacy (%)") +
  scale_y_continuous(breaks = c(-100,-80,-60,-40,-20, 0,20,40,60,80,100), limits = c(-105,100)) +
  theme_classic(base_size = 30) +
  
  theme(axis.text.x=element_text(angle=45,hjust=1),  axis.title.x=element_blank(), legend.position = "none")

# ggsave("./outputs/efficacypost_2024-03-15 v1.png", width = 12, height = 8.625, units = "in")
```


## Combine Pane A (Varroa predictions) and Pane B (day 77 Efficacy)

```{r}
# Old color scheme - DAN

ggarrange(paneA.lab, paneB, ncol = 2, nrow = 1, heights = c(0.80, 1), widths = c(1, 0.80), labels = "AUTO", font.label = list(size = 30)) # common.legend = TRUE heights = 

ggsave("outputs/combo_var_efficacy_C2 2024-12-23 label.png", width = 17.25, height = 8, units = "in")

ggarrange(paneA.nolab, paneB, ncol = 2, nrow = 1, heights = c(0.80, 1), widths = c(1, 1), labels = "AUTO", font.label = list(size = 30)) # common.legend = TRUE heights = 

ggsave("outputs/combo_var_efficacy_C2 2024-12-23 nolabel.png", width = 17.25, height = 8, units = "in")

#### New color scheme - DAN

ggarrange(paneA.BW.lab, paneB.BW, ncol = 2, nrow = 1, heights = c(0.80, 1), widths = c(1, 0.80), labels = "AUTO", font.label = list(size = 30)) # common.legend = TRUE heights = 

ggsave("outputs/combo_var_efficacy_BW 2024-12-23 label.png", width = 17.25, height = 8, units = "in")

ggarrange(paneA.BW.nolab, paneB.BW, ncol = 2, nrow = 1, heights = c(0.80, 1), widths = c(1, 1), labels = "AUTO", font.label = list(size = 30)) # common.legend = TRUE heights = 

ggsave("outputs/combo_var_efficacy_BW 2024-12-23 nolabel.png", width = 17.25, height = 8, units = "in")



```


# Determine how treatment effectiveness depends on days of brood break

```{r}
# A worker brood cell becomes receptive to Varroa on day 8, and emerges on day 21

temp <- temp %>% 
  mutate(age_start_lay = start_layday - 2,
          emerge_day = 21, # Age when worker brood starts emerging
         receptive_day = start_layday + 8,
         days_break = receptive_day - emerge_day,
         completeYN = ifelse(days_break > 0, "Complete", "Not Complete"),
         completeYN = fct_relevel(completeYN, "Not Complete", "Complete"),
         delta_perc_var = ch77_perc_var-ch2_perc_var
         )
```


```{r}
# Basic structure of model formulas:
# End ~ start + length + start:length

temp %>% 
  ggplot(aes(x = ch2_perc_var, y = ch77_perc_var, color = days_break)) +
  geom_point() +
  facet_wrap(~trt_label)

temp %>% 
  ggplot(aes(x = days_break, y = ch77_perc_var-ch2_perc_var)) +
  geom_point() +
  facet_wrap(~trt_label)

temp %>% 
  ggplot(aes(x = completeYN, y = ch77_perc_var-ch2_perc_var)) +
  geom_boxplot() +
  facet_wrap(~trt_label)

temp %>% 
  ggplot(aes(x = delta_perc_var)) +
  geom_histogram()


m5 <- lmer(delta_perc_var ~ days_break + (1|trt_label), data = temp)
m5.0 <- lmer(delta_perc_var ~ 1 + (1|trt_label), data = temp)
# For every additional day of break, there were 0.05 fewer percent Varroa

summary(m5)
anova(m5, m5.0) # There was not a significant influence of brood break length on delta Varroa (P=0.2661)


m6 <- lmer(delta_perc_var ~ completeYN + (1|trt_label), data = temp)
m6.0 <- lmer(delta_perc_var ~ 1 + (1|trt_label), data = temp)

anova(m6, m6.0)

```








