---
title: "Analysis - Varroa"
author: "Dan Aurell"
date: "2024-01-07"
output:   
  html_document: 
    keep_md: yes
---

# Setup

```{r}
library(MASS)
library(tidyverse)
library(emmeans)
library(broom)
library(multcomp)
library(lme4)
library(lubridate)
library(mgcv)
```


# Read in data

```{r}
an_dat <- read.csv("./data_prepared/an_dat_q.csv")

names(an_dat)
```

# Prepare data for Varroa infestation analysis

Exclude any colonies that had EFB or queen issues or were removed due to weakness
```{r}
temp <- an_dat %>% 
  filter(issues == 0) %>% 
  mutate(ch2_wash_bees = ch2_wash_workers + ch2_wash_drones + ch2_wash_queens,
         ch77_wash_bees = ch77_wash_workers + ch77_wash_drones + ch77_wash_queens,
         chpost_wash_bees = chpost_wash_workers + chpost_wash_drones + chpost_wash_queens,
         ch2_perc_var = 100*ch2_var_total/ch2_wash_bees,
         ch77_perc_var = 100*ch77_var_total/ch77_wash_bees,
         chpost_perc_var = 100*chpost_var_total/chpost_wash_bees
         )

# 117 of 177 colonies still in trial on d77: 66% of colonies. Not too bad.

names(temp)

unique(temp$trt_label)

temp <- temp %>%
  mutate(trt_label = fct_relevel(trt_label, "Control", "Apivar", "Amitraz E.C.", "OA Dribble",  "5x OA Dribble","OA Vapor", "HopGuard"))

# calculate the number of days between day 2 sampling and post sampling
temp <- temp %>% 
  mutate(
    num_days = mdy(chpost_date) - mdy(ch2_date),
    days_numeric = as.numeric(num_days)
  )

mean(unique(temp$days_numeric)) #102.75 ; round to 100

# pattern <- "date"
#   grep(pattern, names(temp), value=TRUE)  


```

# Assess were there differences in Varroa load between treatments?

## Describe levels on day 2, 77, post
```{r}
temp %>% 
  summarise(mean = mean(ch2_perc_var))

temp %>% 
  summarise(mean = mean(ch77_perc_var))

temp %>% 
  summarise(mean = mean(chpost_perc_var))

temp %>%
  group_by(trt_label) %>% 
    summarise(mean2 = mean(ch2_perc_var),
              mean77 = mean(ch77_perc_var),
              meanpost = mean(chpost_perc_var)
              )
```


## Make long version of Varroa data
```{r}
temp_piv <- temp %>% 
  pivot_longer(c(ch2_perc_var, ch77_perc_var, chpost_perc_var), names_to = "day", values_to = "perc_var") 

temp_piv <- temp_piv %>%
  mutate(trt_label = fct_relevel(trt_label, "Control", "Apivar", "Amitraz E.C.", "OA Dribble",  "5x OA Dribble","OA Vapor", "HopGuard"))

# Rename "day" values to 2, 77, post

temp_piv <- temp_piv %>% 
  mutate(day = if_else(day == "ch2_perc_var", "2", 
                       if_else(day == "ch77_perc_var", "77", 
                               if_else(day == "chpost_perc_var", "Post", NA))))

```

## Determine how many were above/below 3% threshold for each treatment each day

```{r}
temp_piv <- temp_piv %>%
  mutate(above = if_else(perc_var >= 3, 1, 0))

# This above/below information is now saved in a df
above_df <- temp_piv %>% 
  group_by(day, trt_label) %>% 
  summarize(n=n(),
            above_num = sum(above),
            above_p = above_num/n,)

```






## Graph differences on day 2, 77, post



Graph it
```{r}
temp_piv %>% 
  ggplot(aes(x=trt_label, y = perc_var, fill = trt_label)) +
  geom_boxplot() +
  ylab("Varroa infestation of\nadult bees (%)") +
  facet_wrap(~day) +
  theme_classic(base_size = 30) +
  theme(axis.text.x=element_text(angle=45,hjust=1), axis.title.x=element_blank(), legend.position = "none")

ggsave("./outputs/supp_varroa_daily v1.png", width = 12, height = 8.625, units = "in")
```

## Model pre-treatment differences

```{r}

## Including only non-issues colonies

m.day2 <- glm.nb(ch2_var_total~trt_label, data = temp)
summary(m.day2)

m.day2.0 <- glm.nb(ch2_var_total~1, data = temp)
summary(m.day2.0)

anova(m.day2, m.day2.0) # Pre-treatment differences were not significant (LR Stat 6 = 4.023; P=0.674)

## Qualitatively same result if include all colonies
m.day2 <- glm.nb(ch2_var_total~trt_label, data = an_dat)
summary(m.day2)

m.day2.0 <- glm.nb(ch2_var_total~1, data = an_dat)
summary(m.day2.0)

anova(m.day2, m.day2.0) # Pre-treatment differences were not significant (P=0.84)


```
# Post-treatment differences

```{r}
an_dat %>% 
  ggplot(aes(x=trt_label, y = ch77_var_total)) +
  geom_boxplot()
# Looks like t

m.day2 <- glm.nb(ch2_var_total~trt_label, data = an_dat)
summary(m.day2)

m.day2.0 <- glm.nb(ch2_var_total~1, data = an_dat)
summary(m.day2.0)

anova(m.day2, m.day2.0) # Pre-treatment differences were not significant (P=0.84)
```



# Run models for d77 varroa

## I would consider including yard as a random effect since we blocked by yard
This should increase power to detect treatment effects
But it gives a singular fit, so I will not use it as a random effect
- But I think Steury was mentioning just make the model more patient with BOBYA

```{r}
m1 <- glmer.nb(ch77_var_total~trt_label + ch2_var_total + (1|new_yard), data=temp)
# singular
summary(m1)
help('isSingular')


m1.1 <- glm.nb(ch77_var_total~trt_label + ch2_var_total, data=temp)
# singular
summary(m1.1)

# SUUUPER SIMILAR RESULTS of coefficients with and without random effect

```






## Correcting for number of bees in sample by including it as an offset

```{r}
m2 <- glmer.nb(ch77_var_total ~ 
                 trt_label + 
                 ch2_perc_var +  # 1 = 1 mite per 100 bees
                 offset(log(ch77_wash_bees)) +
                 (1|new_yard), 
               data=temp)
summary(m2)
# Singular fit, still

# Dr. Steury comment: Could try it with a BOBYQA optimizer...
# ... see if get same estimates of effect
# If same estimates of effect, can use this one, even if it has a singular fit

# Omitting random effect
m3 <- glm.nb(ch77_var_total ~ 
                 trt_label + 
                 ch2_perc_var +
                 offset(log(ch77_wash_bees)), 
               data=temp)
# summary(m3)
summary(m3)

m0 <- glm.nb(ch77_var_total ~ 
                 # trt_label + 
                 ch2_perc_var + 
                 offset(log(ch77_wash_bees)), 
               data=temp)
summary(m0)

anova(m3, m0) # Significant effect of trt

summary(m3) # Very similar estimates of effect


```



## How did the Varroa infestations vary between treatments on day 77?
Used to make statistical comparisons between treatment groups

```{r}
# Make a priori contrasts between treatments
K <- contrMat(table(temp$trt_label), type="Tukey")
Ksub <- rbind(K[c(3,4,5,6,8,9,10,11,16,17),])

summary(glht(m3, linfct = mcp(trt_label = Ksub)),
        test = adjusted("single-step")) # Used for p-values for ESA poster

# Approach varies journal to journal
# One thing I might do is just run all post-hoc comparisons
  # If qualitatively the same results as my initial approach I could just use those
  # If I do it this way, I will need to justify why I'm selecting some comparisons in particular...

```

# Run models for post-honey Varroa
The 4 experimental blocks had a staggered start, but ended at the same time (due to the honey harvest )
To account for the different length of time that each yard was in the experiment, I include the number of days as a predictor, and make predictions normalized to ~103 days of experiment length.

```{r}

 pattern <- "bees"
   grep(pattern, names(temp), value=TRUE) 

m4 <- glm.nb(chpost_var_total ~ 
                 trt_label + 
                 ch2_perc_var + 
               days_numeric +
                 offset(log(chpost_wash_bees)), 
               data=temp)

summary(m4)





m5 <- glm.nb(chpost_var_total ~ 
                 trt_label + 
                 ch2_perc_var + 
               # days_numeric +
                 offset(log(chpost_wash_bees)), 
               data=temp)

summary(m5)

anova(m4, m5)
# More complex model favored p = 0.003; go with m4
```

## How did the Varroa infestations vary between treatments on day 103?
Used to make statistical comparisons between treatment groups

```{r}
# Make a priori contrasts between treatments
K <- contrMat(table(temp$trt_label), type="Tukey")
Ksub <- rbind(K[c(3,4,5,6,8,9,10,11,16,17),])

summary(glht(m4, linfct = mcp(trt_label = Ksub)),
        test = adjusted("single-step"))

```


# Summarize results as efficacy
- And get confidence intervals around efficacy

## Day 77 Efficacy

```{r}


tidy(m3)

eff77 <- tidy(m3)

eff77_CI <- as.data.frame(confint(m3))
eff77 <- cbind(eff77, eff77_CI)
eff77 <- eff77 %>% 
  rename(LCL = "2.5 %", UCL = "97.5 %")



eff77 <- eff77 %>% 
  mutate(times_as = exp(estimate),
         efficacy77 = 100*(1-times_as),
         
         times_as_UCL = exp(UCL),
         efficacy77_LCL = 100*(1-times_as_UCL),
         
         times_as_LCL = exp(LCL),
         efficacy77_UCL = 100*(1-times_as_LCL)
         )

eff77 <- eff77[2:7,]
```

## Day 103 Efficacy

```{r}

effpost <- tidy(m4)

effpost_CI <- as.data.frame(confint(m4))
effpost <- cbind(effpost, effpost_CI)
effpost <- effpost %>% 
  rename(LCL = "2.5 %", UCL = "97.5 %")


effpost <- effpost %>% 
  mutate(times_as = exp(estimate),
         efficacypost = 100*(1-times_as),
         
         times_as_UCL = exp(UCL),
         efficacypost_LCL = 100*(1-times_as_UCL),
         
         times_as_LCL = exp(LCL),
         efficacypost_UCL = 100*(1-times_as_LCL)
         )

effpost <- effpost[2:7,]

```


## Plot d77 efficacy

```{r}
eff77 %>% 
  ggplot(aes(y = efficacy77, x = term), fill = "#031C38", alpha = 0.6) + 
  geom_bar(stat = "identity", color = "black") +
  geom_errorbar(aes(ymin=efficacy77_LCL, ymax=efficacy77_UCL), width=0.5) +
  
  ylab("Efficacy (%)") +
  scale_y_continuous(breaks = c(0,20,40,60,80,100)) +
  theme_classic() + # base_size = 30
  
  theme(axis.text.x=element_text(angle=45,hjust=1),  axis.title.x=element_blank())
  
```
## Plot d103 efficacy

```{r}
effpost %>% 
  ggplot(aes(y = efficacypost, x = term), fill = "#031C38", alpha = 0.6) + 
  geom_bar(stat = "identity", color = "black") +
  geom_errorbar(aes(ymin=efficacypost_LCL, ymax=efficacypost_UCL), width=0.5) +
  
  ylab("Efficacy (%)") +
  # scale_y_continuous(breaks = c(0,20,40,60,80,100)) +
  theme_classic() + # base_size = 30
  
  theme(axis.text.x=element_text(angle=45,hjust=1),  axis.title.x=element_blank())
  
```





Get predictions for day 77
```{r}
pred_77 <- emmip(m3, ~ trt_label, type = "response", CIs = TRUE, plotit = FALSE, offset = log(100))
# this is based on the non-included variable (Varroa infestation on day 2) being set to its mean... 
# Predictions are now set to the value of offset = 100... (mites per 100 bees)

pred_77$day <- 77

# Check this against the actual means per group
temp %>% 
  group_by(trt_label) %>% 
  summarise(varroa = mean(ch77_perc_var))
# Matches closely



```

Get predictions for day 103

```{r}
pred_post <- emmip(m4, ~ trt_label, type = "response", CIs = TRUE, plotit = FALSE, offset = log(100))

pred_post$day <- 103

```


```{r}
# Combine day 2, 77, 103 into one plottable data frame

mean(temp$ch2_perc_var) # 1.42 Varroa per 100 bees

pred_pre <- data.frame(pred_77$trt_label, yvar = c(rep(1.42, 7)), day = c(rep(2, 7)))
pred_pre <- pred_pre %>% 
  rename(trt_label = pred_77.trt_label)

  
  
# eff <- data.frame(eff77$term, eff77$efficacy77, effpost$efficacypost[1:8])
  
  

pred <- bind_rows(pred_pre, pred_77, pred_post)
```


# Graph Varroa infestation predictions

```{r}
pred %>% 
  ggplot(mapping = aes(y = yvar, x = day, colour = trt_label)) +
  geom_hline(yintercept = 3, linetype = 2, colour = "red") +
  geom_line(size = 1.5) +
  geom_point(size = 3) +

  scale_colour_manual(
  limits = c("Control", "OA Vapor", "HopGuard", "OA Dribble", "Apivar", "5x OA Dribble", "Amitraz E.C."),
  labels = c("Control", "OA Vapor", "HopGuard", "OA Dribble", "Apivar", "5x OA Dribble", "Amitraz E.C."),
  values = c("#F8766D", "#C49A00", "#53B400", "#00C094", "#00B6EB", "#A58AFF", "#FB61D7"), 
  aesthetics = "color") +
  
  scale_x_continuous(breaks = c(2, 77, 103)) +
  # ylim(0,8) +
  
  ylab("Varroa infestation of\nadult bees (%)") +
  xlab("Experiment day")
  
  # theme_classic(base_size = 30)
```












