---
title: "Analysis - Varroa"
author: "Dan Aurell"
date: "2024-01-07"
output:   
  html_document: 
    keep_md: yes
---

# Setup

```{r}
library(MASS)
library(tidyverse)
library(emmeans)
library(broom)
library(multcomp)
library(lme4)
library(lubridate)
library(mgcv)
```



# Read in data

```{r}
an_dat <- read.csv("./data_prepared/an_dat_q.csv")

names(an_dat)
```

# Prepare data for Varroa infestation analysis

Exclude any colonies that had EFB or queen issues or were removed due to weakness
```{r}
temp <- an_dat %>% 
  filter(issues == 0) %>% 
  mutate(ch2_wash_bees = ch2_wash_workers + ch2_wash_drones + ch2_wash_queens,
         ch77_wash_bees = ch77_wash_workers + ch77_wash_drones + ch77_wash_queens,
         chpost_wash_bees = chpost_wash_workers + chpost_wash_drones + chpost_wash_queens,
         ch2_perc_var = 100*ch2_var_total/ch2_wash_bees,
         ch77_perc_var = 100*ch77_var_total/ch77_wash_bees,
         chpost_perc_var = 100*chpost_var_total/chpost_wash_bees
         )

# 117 of 177 colonies still in trial on d77: 66% of colonies. Not too bad.

names(temp)

unique(temp$trt_label)

temp <- temp %>%
  mutate(trt_label = fct_relevel(trt_label, "Control", "Apivar", "Amitraz E.C.", "OA Dribble",  "5x OA Dribble","OA Vapor", "HopGuard"))

# calculate the number of days between day 2 sampling and post sampling
temp <- temp %>% 
  mutate(
    num_days = mdy(chpost_date) - mdy(ch2_date),
    days_numeric = as.numeric(num_days)
  )

mean(unique(temp$days_numeric)) #102.75 ; round to 100

# pattern <- "date"
#   grep(pattern, names(temp), value=TRUE)  


```

# Assess were there pre-treatment and post-treatment differences in Varroa load?

## Describe levels on day 2 and 77
```{r}
temp %>% 
  summarise(mean = mean(ch2_perc_var))

temp %>% 
  summarise(mean = mean(ch77_perc_var))

temp %>% 
  summarise(mean = mean(chpost_perc_var))

temp %>%
  filter(trt_label == "Control")
    summarise(mean = mean(ch2_perc_var))

temp %>% 
  summarise(mean = mean(ch77_perc_var))

temp %>% 
  summarise(mean = mean(chpost_perc_var))

```



## Graph differences

```{r}
an_dat %>% 
  ggplot(aes(x=trt_label, y = ch2_var_total)) +
  geom_boxplot()
# Doesn't look like it at first glance when all data are included

an_dat %>% 
  filter(issues == 0) %>% 
  ggplot(aes(x=trt_label, y = ch2_var_total)) +
  geom_boxplot()
# But maybe HopGuard is high when we exclude colonies with issues



an_dat %>% 
  ggplot(aes(x=trt_label, y = ch77_var_total)) +
  geom_boxplot()

an_dat %>% 
  filter(issues == 0) %>% 
  ggplot(aes(x=trt_label, y = ch77_var_total)) +
  geom_boxplot()

# The stories are really similar whether you include the colonies with "issues" or not, but this just makes sense since the ones with issues had already been removed before day 77


```




```{r}
## Including all colonies

m.day2 <- glm.nb(ch2_var_total~trt_label, data = an_dat)
summary(m.day2)

m.day2.0 <- glm.nb(ch2_var_total~1, data = an_dat)
summary(m.day2.0)

anova(m.day2, m.day2.0) # Pre-treatment differences were not significant (P=0.84)


## Including only non-issues colonies

m.day2 <- glm.nb(ch2_var_total~trt_label, data = temp)
summary(m.day2)

m.day2.0 <- glm.nb(ch2_var_total~1, data = temp)
summary(m.day2.0)

anova(m.day2, m.day2.0) # Pre-treatment differences were not significant (P=0.67)
```
# Post-treatment differences

```{r}
an_dat %>% 
  ggplot(aes(x=trt_label, y = ch77_var_total)) +
  geom_boxplot()
# Looks like t

m.day2 <- glm.nb(ch2_var_total~trt_label, data = an_dat)
summary(m.day2)

m.day2.0 <- glm.nb(ch2_var_total~1, data = an_dat)
summary(m.day2.0)

anova(m.day2, m.day2.0) # Pre-treatment differences were not significant (P=0.84)
```



# Run models for d77 varroa

## I would consider including yard as a random effect since we blocked by yard
This should increase power to detect treatment effects
But it gives a singular fit, so I will not use it as a random effect
- But I think Steury was mentioning just make the model more patient with BOBYA

```{r}
m1 <- glmer.nb(ch77_var_total~trt_label + ch2_var_total + (1|new_yard), data=temp)
# singular
summary(m1)
help('isSingular')


m1.1 <- glm.nb(ch77_var_total~trt_label + ch2_var_total, data=temp)
# singular
summary(m1.1)

# SUUUPER SIMILAR RESULTS of coefficients with and without random effect

```






## Correcting for number of bees in sample by including it as an offset

```{r}
m2 <- glmer.nb(ch77_var_total ~ 
                 trt_label + 
                 ch2_perc_var +  # 1 = 1 mite per 100 bees
                 offset(log(ch77_wash_bees)) +
                 (1|new_yard), 
               data=temp)
summary(m2)
# Singular fit, still

# Dr. Steury comment: Could try it with a BOBYQA optimizer...
# ... see if get same estimates of effect
# If same estimates of effect, can use this one, even if it has a singular fit

# Omitting random effect
m3 <- glm.nb(ch77_var_total ~ 
                 trt_label + 
                 ch2_perc_var +
                 offset(log(ch77_wash_bees)), 
               data=temp)
# summary(m3)
summary(m3)

m0 <- glm.nb(ch77_var_total ~ 
                 # trt_label + 
                 ch2_perc_var + 
                 offset(log(ch77_wash_bees)), 
               data=temp)
summary(m0)

anova(m3, m0) # Significant effect of trt

summary(m3) # Very similar estimates of effect


```



## How did the Varroa infestations vary between treatments on day 77?
Used to make statistical comparisons between treatment groups

```{r}
# Make a priori contrasts between treatments
K <- contrMat(table(temp$trt_label), type="Tukey")
Ksub <- rbind(K[c(3,4,5,6,8,9,10,11,16,17),])

summary(glht(m3, linfct = mcp(trt_label = Ksub)),
        test = adjusted("single-step")) # Used for p-values for ESA poster

# Approach varies journal to journal
# One thing I might do is just run all post-hoc comparisons
  # If qualitatively the same results as my initial approach I could just use those
  # If I do it this way, I will need to justify why I'm selecting some comparisons in particular...

```

# Run models for post-honey Varroa
The 4 experimental blocks had a staggered start, but ended at the same time (due to the honey harvest )
To account for the different length of time that each yard was in the experiment, I include the number of days as a predictor, and make predictions normalized to ~103 days of experiment length.

```{r}

 pattern <- "bees"
   grep(pattern, names(temp), value=TRUE) 

m4 <- glm.nb(chpost_var_total ~ 
                 trt_label + 
                 ch2_perc_var + 
               days_numeric +
                 offset(log(chpost_wash_bees)), 
               data=temp)

summary(m4)





m5 <- glm.nb(chpost_var_total ~ 
                 trt_label + 
                 ch2_perc_var + 
               # days_numeric +
                 offset(log(chpost_wash_bees)), 
               data=temp)

summary(m5)

anova(m4, m5)
# More complex model favored p = 0.003; go with m4
```

## How did the Varroa infestations vary between treatments on day 103?
Used to make statistical comparisons between treatment groups

```{r}
# Make a priori contrasts between treatments
K <- contrMat(table(temp$trt_label), type="Tukey")
Ksub <- rbind(K[c(3,4,5,6,8,9,10,11,16,17),])

summary(glht(m4, linfct = mcp(trt_label = Ksub)),
        test = adjusted("single-step"))

```


# Summarize results as efficacy
- And get confidence intervals around efficacy

## Day 77 Efficacy

```{r}


tidy(m3)

eff77 <- tidy(m3)

eff77_CI <- as.data.frame(confint(m3))
eff77 <- cbind(eff77, eff77_CI)
eff77 <- eff77 %>% 
  rename(LCL = "2.5 %", UCL = "97.5 %")



eff77 <- eff77 %>% 
  mutate(times_as = exp(estimate),
         efficacy77 = 100*(1-times_as),
         
         times_as_UCL = exp(UCL),
         efficacy77_LCL = 100*(1-times_as_UCL),
         
         times_as_LCL = exp(LCL),
         efficacy77_UCL = 100*(1-times_as_LCL)
         )

eff77 <- eff77[2:7,]
```

## Day 103 Efficacy

```{r}

effpost <- tidy(m4)

effpost_CI <- as.data.frame(confint(m4))
effpost <- cbind(effpost, effpost_CI)
effpost <- effpost %>% 
  rename(LCL = "2.5 %", UCL = "97.5 %")


effpost <- effpost %>% 
  mutate(times_as = exp(estimate),
         efficacypost = 100*(1-times_as),
         
         times_as_UCL = exp(UCL),
         efficacypost_LCL = 100*(1-times_as_UCL),
         
         times_as_LCL = exp(LCL),
         efficacypost_UCL = 100*(1-times_as_LCL)
         )

effpost <- effpost[2:7,]

```


## Plot d77 efficacy

```{r}
eff77 %>% 
  ggplot(aes(y = efficacy77, x = term), fill = "#031C38", alpha = 0.6) + 
  geom_bar(stat = "identity", color = "black") +
  geom_errorbar(aes(ymin=efficacy77_LCL, ymax=efficacy77_UCL), width=0.5) +
  
  ylab("Efficacy (%)") +
  scale_y_continuous(breaks = c(0,20,40,60,80,100)) +
  theme_classic() + # base_size = 30
  
  theme(axis.text.x=element_text(angle=45,hjust=1),  axis.title.x=element_blank())
  
```
## Plot d103 efficacy

```{r}
effpost %>% 
  ggplot(aes(y = efficacypost, x = term), fill = "#031C38", alpha = 0.6) + 
  geom_bar(stat = "identity", color = "black") +
  geom_errorbar(aes(ymin=efficacypost_LCL, ymax=efficacypost_UCL), width=0.5) +
  
  ylab("Efficacy (%)") +
  # scale_y_continuous(breaks = c(0,20,40,60,80,100)) +
  theme_classic() + # base_size = 30
  
  theme(axis.text.x=element_text(angle=45,hjust=1),  axis.title.x=element_blank())
  
```





Get predictions for day 77
```{r}
pred_77 <- emmip(m3, ~ trt_label, type = "response", CIs = TRUE, plotit = FALSE, offset = log(100))
# this is based on the non-included variable (Varroa infestation on day 2) being set to its mean... 
# Predictions are now set to the value of offset = 100... (mites per 100 bees)

pred_77$day <- 77

# Check this against the actual means per group
temp %>% 
  group_by(trt_label) %>% 
  summarise(varroa = mean(ch77_perc_var))
# Matches closely



```

Get predictions for day 103

```{r}
pred_post <- emmip(m4, ~ trt_label, type = "response", CIs = TRUE, plotit = FALSE, offset = log(100))

pred_post$day <- 103

```


```{r}
# Combine day 2, 77, 103 into one plottable data frame

mean(temp$ch2_perc_var) # 1.42 Varroa per 100 bees

pred_pre <- data.frame(pred_77$trt_label, yvar = c(rep(1.42, 7)), day = c(rep(2, 7)))
pred_pre <- pred_pre %>% 
  rename(trt_label = pred_77.trt_label)

  
  
# eff <- data.frame(eff77$term, eff77$efficacy77, effpost$efficacypost[1:8])
  
  

pred <- bind_rows(pred_pre, pred_77, pred_post)
```


# Graph Varroa infestation predictions

```{r}
pred %>% 
  ggplot(mapping = aes(y = yvar, x = day, colour = trt_label)) +
  geom_hline(yintercept = 3, linetype = 2, colour = "red") +
  geom_line(size = 1.5) +
  geom_point(size = 3) +

  scale_colour_manual(
  limits = c("Control", "OA Vapor", "HopGuard", "OA Dribble", "Apivar", "5x OA Dribble", "Amitraz E.C."),
  labels = c("Control", "OA Vapor", "HopGuard", "OA Dribble", "Apivar", "5x OA Dribble", "Amitraz E.C."),
  values = c("#F8766D", "#C49A00", "#53B400", "#00C094", "#00B6EB", "#A58AFF", "#FB61D7"), 
  aesthetics = "color") +
  
  scale_x_continuous(breaks = c(2, 77, 103)) +
  # ylim(0,8) +
  
  ylab("Varroa / 100 bees") +
  xlab("Experiment day")
  
  # theme_classic(base_size = 30)
```












