---
title: "Analysis - Varroa"
author: "Dan Aurell"
date: "2024-01-07"
output:   
  html_document: 
    keep_md: yes
---

# Setup

```{r}

library(MASS)
library(tidyverse)
library(emmeans)
library(broom)
library(multcomp)
library(lme4)
library(lubridate)
library(mgcv)
```

# Read in data

```{r}
an_dat <- read.csv("./data_prepared/an_dat.csv")

names(an_dat)
```

# Prepare data for Varroa infestation analysis

Exclude any colonies that had EFB or queen issues or were removed due to weakness
```{r}
temp <- an_dat %>% 
  filter(issues == 0) %>% 
  mutate(ch2_wash_bees = ch2_wash_workers + ch2_wash_drones + ch2_wash_queens,
         ch77_wash_bees = ch77_wash_workers + ch77_wash_drones + ch77_wash_queens,
         chpost_wash_bees = chpost_wash_workers + chpost_wash_drones + chpost_wash_queens,
         ch2_perc_var = 100*ch2_var_total/ch2_wash_bees,
         ch77_perc_var = 100*ch77_var_total/ch77_wash_bees,
         chpost_perc_var = 100*chpost_var_total/chpost_wash_bees
         )

# 117 of 177 colonies still in trial on d77: 66% of colonies. Not too bad.

names(temp)

unique(temp$trt_label)

temp <- temp %>%
  mutate(trt_label = fct_relevel(trt_label, "Control", "Apivar", "Amitraz E.C.", "OA Dribble",  "5x OA Dribble","OA Vapor", "HopGuard"))

# calculate the number of days between day 2 sampling and post sampling
temp <- temp %>% 
  mutate(
    num_days = mdy(chpost_date) - mdy(ch2_date),
    days_numeric = as.numeric(num_days)
  )

unique(temp$days_numeric) # 95 to 113 days
mean(unique(temp$days_numeric)) #102.75 ; round to 100

# pattern <- "date"
#   grep(pattern, names(temp), value=TRUE)  


```

# Assess were there differences in Varroa load between treatments?

## Describe arithmetic means on day 2, 77, post
```{r}
temp %>% 
  summarise(mean = mean(ch2_perc_var))

temp %>% 
  summarise(mean = mean(ch77_perc_var))

temp %>% 
  summarise(mean = mean(chpost_perc_var))

temp %>%
  group_by(trt_label) %>% 
    summarise(mean2 = mean(ch2_perc_var),
              mean77 = mean(ch77_perc_var),
              meanpost = mean(chpost_perc_var)
              )


# Greatest numerical difference between 5x OA dribble and HopGuard

# 2.0474221/0.8906983 = 2.3 times as many mites in the HopGuard group as in the 5x OA dribble. This justifies correcting for it statistically.
```


## Make long version of Varroa data
```{r}
temp_piv <- temp %>% 
  pivot_longer(c(ch2_perc_var, ch77_perc_var, chpost_perc_var), names_to = "day", values_to = "perc_var") 

temp_piv <- temp_piv %>%
  mutate(trt_label = fct_relevel(trt_label, "Control", "Apivar", "Amitraz E.C.", "OA Dribble",  "5x OA Dribble","OA Vapor", "HopGuard"))

# Rename "day" values to 2, 77, post

temp_piv <- temp_piv %>% 
  mutate(day = if_else(day == "ch2_perc_var", "2", 
                       if_else(day == "ch77_perc_var", "77", 
                               if_else(day == "chpost_perc_var", "Post", NA))))

```

## Determine how many were above/below 3% threshold for each treatment each day

```{r}
temp_piv <- temp_piv %>%
  mutate(above = if_else(perc_var >= 3, 1, 0))

# This above/below information is now saved in a df
above_df <- temp_piv %>% 
  group_by(day, trt_label) %>% 
  summarize(n=n(),
            above_num = sum(above),
            above_perc = 100*above_num/n,)

```

## Graph Varroa levels on day 2, 77, post

Graph it
```{r}
temp_piv %>% 
  ggplot(aes(x=trt_label, y = perc_var, fill = trt_label)) +
  geom_boxplot() +
  ylab("Varroa infestation of\nadult bees (%)") +
  facet_wrap(~day) +
  theme_classic(base_size = 30) +
  theme(axis.text.x=element_text(angle=45,hjust=1), axis.title.x=element_blank(), legend.position = "none")

# ggsave("./outputs/supp_varroa_daily v1.png", width = 12, height = 8.625, units = "in")
```

## Assess pre-treatment differences statistically

```{r}

## Including only non-issues colonies

m.day2 <- glm.nb(ch2_var_total~trt_label, data = temp)
summary(m.day2)

m.day2.0 <- glm.nb(ch2_var_total~1, data = temp)
summary(m.day2.0)

anova(m.day2, m.day2.0) # Pre-treatment differences were not significant (ChiSq 6 = 4.023; P=0.674)

## Qualitatively same result if include all colonies
m.day2 <- glm.nb(ch2_var_total~trt_label, data = an_dat)
summary(m.day2)

m.day2.0 <- glm.nb(ch2_var_total~1, data = an_dat)
summary(m.day2.0)

anova(m.day2, m.day2.0) # Pre-treatment differences were not significant (P=0.84)


```
# Post-treatment differences

```{r}
an_dat %>% 
  ggplot(aes(x=trt_label, y = ch77_var_total)) +
  geom_boxplot()
# Looks like t

m.day77 <- glm.nb(ch77_var_total~trt_label, data = an_dat)
summary(m.day2)

m.day77.0 <- glm.nb(ch77_var_total~1, data = an_dat)
summary(m.day2.0)

anova(m.day77, m.day77.0) # Post-treatment differences were significant (ChiSq 6 = 28.9; P=6.30e-05)
```



# Run models for d77 varroa

## I would consider including yard as a random effect since we blocked by yard
This should increase power to detect treatment effects
But it gives a singular fit, so I will not use it as a random effect
- But I think Steury was mentioning just make the model more patient with BOBYQA

```{r}
# m1 <- glmer.nb(ch77_var_total~trt_label + ch2_var_total + (1|new_yard), data=temp)
# singular help('isSingular')
summary(m1)



m1.1 <- glm.nb(ch77_var_total~trt_label + ch2_var_total, data=temp)
# Not singular
summary(m1.1)

# SUUUPER SIMILAR RESULTS of coefficients with and without random effect

```






## Correcting for number of bees in sample by including it as an offset

```{r}
m2 <- glmer.nb(ch77_var_total ~ 
                 trt_label + 
                 ch2_perc_var +  # 1 = 1 mite per 100 bees
                 offset(log(ch77_wash_bees)) +
                 (1|new_yard), 
               data=temp)
summary(m2)
# Singular fit

# Dr. Steury comment: Could try it with a BOBYQA optimizer...
# ... see if get same estimates of effect; Yes, I do.

# If same estimates of effect as the model without random effect, can use this one, even if it has a singular fit
# Yes, the estimates of effect are SO close that I will use the singular-fit model.


m2.0 <- glmer.nb(ch77_var_total ~ 
                 # trt_label + 
                 ch2_perc_var +  # 1 = 1 mite per 100 bees
                 offset(log(ch77_wash_bees)) +
                 (1|new_yard), 
               data=temp)
summary(m2.0)

anova(m2, m2.0) # There was a significant effect of treatment on day-77 Varroa load (ChiSq 6 = 37.279; P=1.554e-06)

# Switching to BOBYQA optimizer using glmer argument glmerControl(optimizer="bobyqa") still gives singular fit.


# Omitting random effect
m3 <- glm.nb(ch77_var_total ~ 
                 trt_label + 
                 ch2_perc_var +
                 offset(log(ch77_wash_bees)), 
               data=temp)
# summary(m3)
summary(m3)

m0 <- glm.nb(ch77_var_total ~ 
                 # trt_label + 
                 ch2_perc_var + 
                 offset(log(ch77_wash_bees)), 
               data=temp)
summary(m0)

anova(m3, m0) 
# Treatment significantly influenced Varroa load on day 77 (ChiSq 6 = 37.27869, P=1.553757e-06=0.000001553757)




summary(m3) 

summary(m2)
summary(m3)# Very similar estimates of effect

```



## How did the Varroa infestations vary between treatments on day 77?
Used to make statistical comparisons between treatment groups

```{r}
# Make a priori contrasts between treatments
K <- contrMat(table(temp$trt_label), type="Tukey") # all contrasts
Ksub <- rbind(K[c(3,4,5,6,8,9,10,11,16,17),]) # A priori hypotheses
Ksub2 <- rbind(K[c(1:6),]) # Curious what happens if I just compare each treatment to control

summary(glht(m2, linfct = mcp(trt_label = Ksub)),
        test = adjusted("single-step"))

# Steury comment: Approach varies journal to journal
# If I do it this way that I originally planned (a priori contrasts), I will need to justify why I'm selecting some comparisons in particular...

# One thing I might do is just run all post-hoc comparisons
  # If qualitatively the same results as my initial approach I could just use those

summary(glht(m2, linfct = mcp(trt_label = K)),
        test = adjusted("single-step"))
# Yes, this does provide qualitatively the same results. 
# I use this for reporting results

# Curious how much influence the p-value correction has
summary(glht(m2, linfct = mcp(trt_label = K)),
        test = adjusted("none"))

# I want to be conservative and make sure I'm not unnecessarily writing off OA vapor. If I test each treatment against control, is OA still stat similar to control?

summary(glht(m2, linfct = mcp(trt_label = Ksub2)),
        test = adjusted("single-step"))
# Yes, OA is still not distinguishable from control.



``` 

	 Simultaneous Tests for General Linear Hypotheses

Multiple Comparisons of Means: User-defined Contrasts


Fit: glm.nb(formula = ch77_var_total ~ trt_label + ch2_perc_var + 
    offset(log(ch77_wash_bees)), data = temp, init.theta = 0.8927878706, 
    link = log)

Linear Hypotheses:
                                  Estimate Std. Error z value Pr(>|z|)    
Apivar - Control == 0             -1.62365    0.41092  -3.951  0.00158 ** 
Amitraz E.C. - Control == 0       -2.07619    0.42998  -4.829  < 0.001 ***
OA Dribble - Control == 0         -1.40645    0.39976  -3.518  0.00787 ** 
5x OA Dribble - Control == 0      -1.72365    0.38863  -4.435  < 0.001 ***
OA Vapor - Control == 0           -0.58540    0.39186  -1.494  0.74787    
HopGuard - Control == 0           -1.75298    0.42241  -4.150  < 0.001 ***
Amitraz E.C. - Apivar == 0        -0.45254    0.45102  -1.003  0.95321    
OA Dribble - Apivar == 0           0.21720    0.42238   0.514  0.99866    
5x OA Dribble - Apivar == 0       -0.10000    0.41209  -0.243  0.99998    
OA Vapor - Apivar == 0             1.03825    0.41488   2.503  0.15754    
HopGuard - Apivar == 0            -0.12933    0.44367  -0.292  0.99995    
OA Dribble - Amitraz E.C. == 0     0.66974    0.44044   1.521  0.73176    
5x OA Dribble - Amitraz E.C. == 0  0.35254    0.43205   0.816  0.98334    
OA Vapor - Amitraz E.C. == 0       1.49079    0.43301   3.443  0.01030 *  
HopGuard - Amitraz E.C. == 0       0.32321    0.45953   0.703  0.99241    
5x OA Dribble - OA Dribble == 0   -0.31720    0.40140  -0.790  0.98590    
OA Vapor - OA Dribble == 0         0.82105    0.40348   2.035  0.39126    
HopGuard - OA Dribble == 0        -0.34653    0.43252  -0.801  0.98486    
OA Vapor - 5x OA Dribble == 0      1.13825    0.39380   2.890  0.05869 .  
HopGuard - 5x OA Dribble == 0     -0.02933    0.42551  -0.069  1.00000    
HopGuard - OA Vapor == 0          -1.16758    0.42470  -2.749  0.08585 .  
---

(Adjusted p values reported -- single-step method)


Several natural treatments had lower Varroa levels than the control:
HopGuard was significantly lower than control (z=-4.150, p< 0.001), as was 5x OA Dribble (z=-4.435, p< 0.001) and OA Dribble (z=-3.518, p0.00787). However, the difference between Control and OA Vapor was not significant (z=-1.494, p=0.74787).

The only other significant difference between treatments was that OA Vapor had more Varroa mites than Amitraz E.C. (z=3.443, P=0.01030).


A - Control
AB - OA Vapor
BC - OA dribble
BC- 5x OA dribble
BC - HopGuard
BC - Apivar
C - AEC


# Run models for post-honey Varroa
The 4 experimental blocks had a staggered start, but ended at the same time (due to the honey harvest )
To account for the different length of time that each yard was in the experiment, I include the number of days as a predictor, and make predictions normalized to ~103 days of experiment length.

```{r}

 pattern <- "bees"
   grep(pattern, names(temp), value=TRUE) 

m4 <- glm.nb(chpost_var_total ~ 
                 trt_label + 
                 ch2_perc_var + 
               days_numeric +
                 offset(log(chpost_wash_bees)), 
               data=temp)

summary(m4)


m4.0 <- glm.nb(chpost_var_total ~ 
                 trt_label + 
                 ch2_perc_var + 
               # days_numeric +
                 offset(log(chpost_wash_bees)), 
               data=temp)

summary(m4.0)

anova(m4, m4.0)
# Effect of days was significant p = 0.003; go with m4

m4.00 <- glm.nb(chpost_var_total ~ 
                 # trt_label + 
                 ch2_perc_var + 
               days_numeric +
                 offset(log(chpost_wash_bees)), 
               data=temp)

summary(m4.00)
anova(m4, m4.00)
# Effect of treatment was significant on day 103 as well (ChiSq 6 = 29.3; P=5.306197e-05=0.00005306197) 




m5 <- glmer.nb(chpost_var_total ~ 
               trt_label + 
               ch2_perc_var +
               days_numeric + # Very collinear with random effect
               offset(log(chpost_wash_bees)) +
             (1|new_yard), # Very collinear with random effect
             data=temp)
summary(m5)


```

## How did the Varroa infestations vary between treatments on day 103?
Used to make statistical comparisons between treatment groups

```{r}
# Make a priori contrasts between treatments
K <- contrMat(table(temp$trt_label), type="Tukey")
Ksub <- rbind(K[c(3,4,5,6,8,9,10,11,16,17),])

summary(glht(m4, linfct = mcp(trt_label = Ksub)),
        test = adjusted("single-step"))

summary(glht(m4, linfct = mcp(trt_label = K)),
        test = adjusted("single-step"))

```


# Summarize results as efficacy
- And get confidence intervals around efficacy

## Day 77 Efficacy - no random effect of yard included

```{r}
tidy(m3)
summary(m3)

eff77_3 <- tidy(m3)

eff77_3CI <- as.data.frame(confint(m3))
eff77_3b <- cbind(eff77_3, eff77_3CI)
eff77_3b <- eff77_3b %>% 
  rename(LCL = "2.5 %", UCL = "97.5 %")



eff77_3b <- eff77_3b %>% 
  mutate(times_as = exp(estimate),
         efficacy77 = 100*(1-times_as),
         
         times_as_UCL = exp(UCL),
         efficacy77_LCL = 100*(1-times_as_UCL),
         
         times_as_LCL = exp(LCL),
         efficacy77_UCL = 100*(1-times_as_LCL)
         )

eff77_3b <- eff77_3b[2:7,]


# Make more pleasant treatment names in a trt_label column
eff77_3b <- eff77_3b %>% 
  mutate(
    trt_label = if_else(term == "trt_labelApivar", "Apivar",
                    if_else(term == "trt_labelOA Dribble", "OA Dribble", 
                            if_else(term == "trt_label5x OA Dribble", "5x OA Dribble",
                                   if_else(term == "trt_labelOA Vapor", "OA Vapor",
                                           if_else(term == "trt_labelHopGuard", "HopGuard",
                                                   if_else(term == "trt_labelAmitraz E.C.", "Amitraz E.C.", "NA"
                                                           )
                                                   ))))))
# Add a trt_type column
eff77_3b <- eff77_3b %>% 
  mutate(
    trt_type = if_else(trt_label == "Apivar", "Synthetic",
                    if_else(trt_label == "OA Dribble", "Natural", 
                            if_else(trt_label == "5x OA Dribble", "Natural",
                                   if_else(trt_label == "OA Vapor", "Natural",
                                           if_else(trt_label == "HopGuard", "Natural",
                                                   if_else(trt_label == "Amitraz E.C.", "Synthetic", "NA"
                                                           )
                                                   ))))))

eff77_3b <- eff77_3b %>% 
  mutate(
    trt_label = fct_relevel(trt_label, "Apivar", "Amitraz E.C.", "OA Dribble", "5x OA Dribble", "OA Vapor", "HopGuard")
  )

```


## Day 77 Efficacy - GLMM with random effect of yard included

```{r}
library(broom.mixed)

tidy(m2)

eff77_2 <- tidy(m2)
eff77_2CI <- as.data.frame(confint(m2))

eff77_2 <- eff77_2[2:7,]
eff77_2CI <- eff77_2CI[3:8,]


eff77_2b <- cbind(eff77_2, eff77_2CI)
eff77_2b <- eff77_2b %>% 
  rename(LCL = "2.5 %", UCL = "97.5 %")



eff77_2b <- eff77_2b %>% 
  mutate(times_as = exp(estimate),
         efficacy77 = 100*(1-times_as),
         
         times_as_UCL = exp(UCL),
         efficacy77_LCL = 100*(1-times_as_UCL),
         
         times_as_LCL = exp(LCL),
         efficacy77_UCL = 100*(1-times_as_LCL)
         )


# Make more pleasant treatment names in a trt_label column
eff77_2b <- eff77_2b %>% 
  mutate(
    trt_label = if_else(term == "trt_labelApivar", "Apivar",
                    if_else(term == "trt_labelOA Dribble", "OA Dribble", 
                            if_else(term == "trt_label5x OA Dribble", "5x OA Dribble",
                                   if_else(term == "trt_labelOA Vapor", "OA Vapor",
                                           if_else(term == "trt_labelHopGuard", "HopGuard",
                                                   if_else(term == "trt_labelAmitraz E.C.", "Amitraz E.C.", "NA"
                                                           )
                                                   ))))))
# Add a trt_type column
eff77_2b <- eff77_2b %>% 
  mutate(
    trt_type = if_else(trt_label == "Apivar", "Synthetic",
                    if_else(trt_label == "OA Dribble", "Natural", 
                            if_else(trt_label == "5x OA Dribble", "Natural",
                                   if_else(trt_label == "OA Vapor", "Natural",
                                           if_else(trt_label == "HopGuard", "Natural",
                                                   if_else(trt_label == "Amitraz E.C.", "Synthetic", "NA"
                                                           )
                                                   ))))))

eff77_2b <- eff77_2b %>% 
  mutate(
    trt_label = fct_relevel(trt_label, "Apivar", "Amitraz E.C.", "OA Dribble", "5x OA Dribble", "OA Vapor", "HopGuard")
  )


```


## Decide what approach to take with Day 77 Efficacy

eff77_2CI <- as.data.frame(confint(m2))
Output:
Computing profile confidence intervals ...
Warning: slightly lower deviances (diff=-5.34328e-12) detected
Warning: Last two rows have identical or NA .zeta values: using minstep
Warning: slightly lower deviances (diff=-2.87628e-11) detected
Warning: non-monotonic profile for .sig01
Warning: unexpected decrease in profile: using minstep
Warning: non-monotonic profile for (Intercept)
Warning: bad spline fit for .sig01: falling back to linear interpolation
Warning: collapsing to unique 'x' values
Warning: bad spline fit for (Intercept): falling back to linear interpolation

Mixed model eff77_2b has near-identical estimates of efficacy and CIs around efficacy as the standard model.
I will choose to use the non-mixed efficacy


## Day 103 Efficacy

Previous code:






```{r}
library(broom)

# Using the non-mixed model

effpost <- tidy(m4)
effpost_CI <- as.data.frame(confint(m4))

effpost <- cbind(effpost, effpost_CI)
effpost <- effpost %>% 
  rename(LCL = "2.5 %", UCL = "97.5 %")


effpost <- effpost %>% 
  mutate(times_as = exp(estimate),
         efficacypost = 100*(1-times_as),
         
         times_as_UCL = exp(UCL),
         efficacypost_LCL = 100*(1-times_as_UCL),
         
         times_as_LCL = exp(LCL),
         efficacypost_UCL = 100*(1-times_as_LCL)
         )

effpost <- effpost[2:7,]

## To Do - clean up this data frame like I did for the eff77 data frame.

# Make more pleasant treatment names in a trt_label column
effpost <- effpost %>% 
  mutate(
    trt_label = if_else(term == "trt_labelApivar", "Apivar",
                    if_else(term == "trt_labelOA Dribble", "OA Dribble", 
                            if_else(term == "trt_label5x OA Dribble", "5x OA Dribble",
                                   if_else(term == "trt_labelOA Vapor", "OA Vapor",
                                           if_else(term == "trt_labelHopGuard", "HopGuard",
                                                   if_else(term == "trt_labelAmitraz E.C.", "Amitraz E.C.", "NA"
                                                           )
                                                   ))))))
# Add a trt_type column
effpost <- effpost %>% 
  mutate(
    trt_type = if_else(trt_label == "Apivar", "Synthetic",
                    if_else(trt_label == "OA Dribble", "Natural", 
                            if_else(trt_label == "5x OA Dribble", "Natural",
                                   if_else(trt_label == "OA Vapor", "Natural",
                                           if_else(trt_label == "HopGuard", "Natural",
                                                   if_else(trt_label == "Amitraz E.C.", "Synthetic", "NA"
                                                           )
                                                   ))))))

effpost <- effpost %>% 
  mutate(
    trt_label = fct_relevel(trt_label, "Apivar", "Amitraz E.C.", "OA Dribble", "5x OA Dribble", "OA Vapor", "HopGuard")
  )



```


## Plot d77 efficacy

```{r}
eff77_3b %>% 
  ggplot(aes(y = efficacy77, x = trt_label, fill = trt_type)) + 
  geom_bar(stat = "identity", color = "black") +
  geom_errorbar(aes(ymin=efficacy77_LCL, ymax=efficacy77_UCL), width=0.5) +
  scale_colour_manual(
    limits = c("Control", "Synthetic", "Natural"),
    labels = c("Control", "Synthetic", "Natural"),
    values = c("white", "#F8766D", "#619CFF"), 
    aesthetics = "fill"
  ) +  
  ylab("Efficacy (%)") +
  scale_y_continuous(breaks = c(-20, 0,20,40,60,80,100)) +
  theme_classic(base_size = 30) +
  
  theme(axis.text.x=element_text(angle=45,hjust=1),  axis.title.x=element_blank(), legend.position = "none")

ggsave("./outputs/efficacy_2024-03-11 v1.png", width = 12, height = 8.625, units = "in")
  
```

## Plot d103 efficacy

New plot
```{r}
effpost %>% 
  ggplot(aes(y = efficacypost, x = trt_label, fill = trt_type), alpha = 0.6) + 
  geom_bar(stat = "identity", color = "black") +
  geom_errorbar(aes(ymin=efficacypost_LCL, ymax=efficacypost_UCL), width=0.5) +
  scale_colour_manual(
    limits = c("Control", "Synthetic", "Natural"),
    labels = c("Control", "Synthetic", "Natural"),
    values = c("white", "#F8766D", "#619CFF"), 
    aesthetics = "fill"
  ) +  
  ylab("Efficacy (%)") +
  scale_y_continuous(breaks = c(-100,-80,-60,-40,-20, 0,20,40,60,80,100), limits = c(-105,100)) +
  theme_classic(base_size = 30) +
  
  theme(axis.text.x=element_text(angle=45,hjust=1),  axis.title.x=element_blank(), legend.position = "none")

ggsave("./outputs/efficacypost_2024-03-15 v1.png", width = 12, height = 8.625, units = "in")
```



Old plot
```{r}
effpost %>% 
  ggplot(aes(y = efficacypost, x = term), fill = "#031C38", alpha = 0.6) + 
  geom_bar(stat = "identity", color = "black") +
  geom_errorbar(aes(ymin=efficacypost_LCL, ymax=efficacypost_UCL), width=0.5) +
  
  ylab("Efficacy (%)") +
  # scale_y_continuous(breaks = c(0,20,40,60,80,100)) +
  theme_classic() + # base_size = 30
  
  theme(axis.text.x=element_text(angle=45,hjust=1),  axis.title.x=element_blank())
  
```





Get predictions for day 77
```{r}
pred_77 <- emmip(m3, ~ trt_label, type = "response", CIs = TRUE, plotit = FALSE, offset = log(100))
# this is based on the non-included variable (Varroa infestation on day 2) being set to its mean... 
# Predictions are now set to the value of offset = 100... (mites per 100 bees)

pred_77$day <- 77

# Check this against the actual means per group
temp %>% 
  group_by(trt_label) %>% 
  summarise(varroa = mean(ch77_perc_var))
# Matches closely



```

Get predictions for day 103

```{r}
pred_post <- emmip(m4, ~ trt_label, type = "response", CIs = TRUE, plotit = FALSE, offset = log(100))

pred_post$day <- 103

```


```{r}
# Combine day 2, 77, 103 into one plottable data frame

mean(temp$ch2_perc_var) # 1.42 Varroa per 100 bees

pred_pre <- data.frame(pred_77$trt_label, yvar = c(rep(1.42, 7)), day = c(rep(2, 7)))
pred_pre <- pred_pre %>% 
  rename(trt_label = pred_77.trt_label)

  
  
# eff <- data.frame(eff77$term, eff77$efficacy77, effpost$efficacypost[1:8])
  
  

pred <- bind_rows(pred_pre, pred_77, pred_post)
```


# Graph Varroa infestation predictions

```{r}

# User-selected color scale

pred %>% 
  ggplot(mapping = aes(y = yvar, x = day, colour = trt_label)) +
  geom_hline(yintercept = 3, linetype = 2, colour = "red") +
  geom_line(size = 1.5) +
  geom_point(size = 3) +

  scale_colour_manual(
  limits = c("Control", "OA Vapor", "HopGuard", "Apivar", "OA Dribble", "5x OA Dribble", "Amitraz E.C."),
  labels = c("Control", "OA Vapor", "HopGuard", "Apivar", "OA Dribble", "5x OA Dribble", "Amitraz E.C."),
  values = c("#F8766D", "#C49A00", "#53B400", "#00C094", "#00B6EB", "#A58AFF", "#FB61D7"), 
  aesthetics = "color") +
  
  scale_x_continuous(breaks = c(2, 77, 103)) +
  ylim(0,7) +
  
  ylab("Varroa infestation of\nadult bees (%)") +
  xlab("Experiment day") +
  theme_classic(base_size = 30) +
  theme(legend.title=element_blank())

ggsave("./outputs/varroa_pred_2024-03-11 v1-manual.png", width = 12, height = 8.625, units = "in")

# Viridis color scale

pred %>% 
  ggplot(mapping = aes(y = yvar, x = day, colour = trt_label)) +
  geom_hline(yintercept = 3, linetype = 2, colour = "red") +
  geom_line(size = 1.5) +
  geom_point(size = 3) +

  scale_colour_viridis_d(
  limits = c("Control", "OA Vapor", "HopGuard", "Apivar", "OA Dribble", "5x OA Dribble", "Amitraz E.C."),
  labels = c("Control", "OA Vapor", "HopGuard", "Apivar", "OA Dribble", "5x OA Dribble", "Amitraz E.C."),
  aesthetics = "color",
  direction = -1) +
  
  scale_x_continuous(breaks = c(2, 77, 103)) +
  scale_y_continuous(breaks = c(0,1,2,3,4,5,6,7), labels = c(0,"","",3,"","",6,""), limits = c(0,7)) +
  
  ylab("Varroa infestation of\nadult bees (%)") +
  xlab("Experiment day") +
  theme_classic(base_size = 30) +
  theme(legend.title=element_blank())

ggsave("./outputs/varroa_pred_2024-03-11 v1-viridis.png", width = 12, height = 8.625, units = "in")



```

# Determine how treatment effectiveness depends on days of brood break

```{r}
# Basic structure of model formulas:
# End ~ start + length + start:length




```






# Make supplememtal tables

## Supp Table S1: emmeans Varroa load predictions

## Supp Table S2: Table of efficacies (d77, d103)

```{r}
eff77_3b %>% 
  select(trt_label, efficacy77, efficacy77_LCL, efficacy77_UCL)


```

## Supp Table S3: Table of linear contrasts between treatments 
- Day 77 single-step corrected
- Day 77 uncorrected
- Day 103 single-step corrected
- Day 103 uncorrected













