---
title: "Analysis - Colony development"
author: "Dan Aurell"
date: "2024-01-07"
output:   
  html_document: 
    keep_md: yes
---

```{r}
library(MASS)
library(tidyverse)
library(lme4)
library(emmeans)
library(ggpubr) # For combo plot
library(scales) # To decide on color scheme
```

# Some useful code bits for plotting

  scale_colour_manual(
    limits = c("Apivar", "AEC", "Apiguard_Api"),
    labels = c("Apivar", "Amitraz EC", "Combination"),
    values = c("black", "black", "#00BA38"), 
    aesthetics = "color"
  ) +


Decide on color scheme
- salmon (synthetic)

```{r}
hue_pal()(3)
show_col(hue_pal()(3))
# Salmon #F8766D
# Blue #619CFF

show_col(hue_pal()(2))
# Salmon same as above
# Teal #00BFC4

# My PowerPoint graph colors
# Light Salmon
# sRGB 244 178 132
  # Hex #F4B284

# Light Blue
# sRGB 181 199 231
  # Hex #B5C7E7
```


# Read in data

```{r}
temp <- read.csv("./data_prepared/an_dat.csv")
```

# Prep data for analysis

```{r}
temp <- temp %>% 
  rename(
    fob_44 = ch44_fob,
    fob_58 = ch58_fob,
    fob_72 = ch72_fob,
    fob_77 = frames_bees,
    fob_105 = post_quick_fob
  ) %>% 
  mutate(
    fob_77 = if_else(weak == 1, 0, fob_77),# Set FOB to 0 for weak colonies
    trt_label = fct_relevel(trt_label, "Control", "Apivar", "Amitraz EC", "OA Dribble", "5x OA Dribble", "OA Vapor", "HopGuard")
    )

# for d77 FOB analysis
temp1 <- temp %>% 
  filter(established == 1,
         QL == 0,
         DL == 0, 
         efb == 0
         ) # Still includes 5 "weak" colonies


# for weight at honey harvest analysis
temp2 <- temp %>% 
  filter(established == 1,
         QL == 0,
         DL == 0, 
         efb == 0,
         weak == 0
         ) %>% 
  filter(is.na(post_issues))
# Excludes 5 "weak" colonies and 10 that had issues on day 103

temp %>% 
  filter(weak == 1) %>% 
  select(fob_77, post_kgs_C)
# Of 5 weak colonies, we have d77 fob data for 1 colony
# we have post kgs data for 1 colony

```



## Analyze day 77 Colony strength

```{r}
m.bees1 <- lm(fob_77~trt_label, data = temp1) 
m.bees1.0 <- lm(fob_77~1, data = temp1) 
anova(m.bees1, m.bees1.0) # No improvement to include a treatment effect (ChiSq 6 = 0.0297, p=0.999)


m.bees2 <- lmer(fob_77~trt_label + (1 | new_yard), data = temp1) # Check whether yard level random effect changes results
m.bees2.0 <- lmer(fob_77~ 1 + (1 | new_yard), data = temp1)
anova(m.bees2, m.bees2.0) # No improvement to include a treatment effect (ChiSq 6 = 0.18, p=0.999)

# Still make predictions based on m2, but no statistical detected effect of treatment

# Send emmeans predictions to its own strength item
strength1 <- emmip(m.bees1, ~trt_label, type = "response", CIs = TRUE, plotit = FALSE)
strength2 <- emmip(m.bees2, ~trt_label, type = "response", CIs = TRUE, plotit = FALSE)
# The results were so similar that I am very happy sticking with the simpler model (no random effect)

emplot_strength <- emmip(m.bees1, ~trt_label, type = "response", CIs = TRUE, plotit = FALSE)
emplot_strength <- emplot_strength %>% 
  rename(
    fob_77 = yvar
  )


emplot_strength <- emplot_strength %>% 
  mutate(
    trt_label = fct_relevel(trt_label, "Control", "Apivar", "Amitraz EC", "OA Dribble", "5x OA Dribble", "OA Vapor", "HopGuard")
  )


```

## Plot day 77 colony strength

```{r}
    # Salmon options #F8766D #F4B284
    # Blue options #619CFF #00BFC4 #B5C7E7

# No color

paneA <- ggplot() +
  geom_boxplot(aes(x = trt_label, y = fob_77), fill = "#031C38", outlier.shape = NA, color = "black", data = temp1) +
  geom_point(aes(x = trt_label, y = fob_77), shape=20, size=10, color="black", fill="black", data = emplot_strength) +
  #xlab("Treatment") +
  ylab("Colony strength\n(frames of bees)") +
  theme_classic( base_size = 30) +
  theme(axis.text.x=element_text(angle=45,hjust=1), axis.title.x=element_blank())

# Color, no treatment names (for vertical layout)

paneA2 <- ggplot() +
  geom_boxplot(aes(x = trt_label, y = fob_77, fill = trt_type), outlier.shape = NA, color = "black", data = temp1) +
    scale_colour_manual(
    limits = c("Control", "Synthetic", "Natural"),
    labels = c("Control", "Synthetic", "Natural"),
    values = c("white", "#F8766D", "#619CFF"), 
    aesthetics = "fill"
  ) +  
  geom_point(aes(x = trt_label, y = fob_77), shape=20, size=10, color="black", fill="black", data = emplot_strength) +
  #xlab("Treatment") +
  ylab("Colony strength\n(frames of bees)") +
  theme_classic( base_size = 30) +
  theme(axis.text.x=element_blank(), axis.title.x=element_blank(), legend.position = "none")

# Color and treatment names (for horizontal layout)

paneA3 <- ggplot() +
  geom_boxplot(aes(x = trt_label, y = fob_77, fill = trt_type), outlier.shape = NA, color = "black", data = temp1) +
    scale_colour_manual(
    limits = c("Control", "Synthetic", "Natural"),
    labels = c("Control", "Synthetic", "Natural"),
    values = c("white", "#F8766D", "#619CFF"), 
    aesthetics = "fill"
  ) +  
  geom_point(aes(x = trt_label, y = fob_77), shape=20, size=10, color="black", fill="black", data = emplot_strength) +
  #xlab("Treatment") +
  ylab("Colony strength\n(frames of bees)") +
  theme_classic( base_size = 30) +
  theme(axis.text.x=element_text(angle=45,hjust=1), axis.title.x=element_blank(), legend.position = "none")



```




## Analyze data for colony weight post honey production

```{r}
m.weight1 <- lm(post_kgs_C~trt_label, data = temp2)
m.weight1.0 <- lm(post_kgs_C~1, data = temp2)
anova(m.weight1.0, m.weight1) # NEW No effect of treatment (F 6 = 0.3604, p=0.9023)

m.weight2 <- lmer(post_kgs_C ~ trt_label + (1 | new_yard), data = temp2)
m.weight2.0 <- lmer(post_kgs_C ~ 1 + (1 | new_yard), data = temp2)
anova(m.weight2, m.weight2.0) # No effect of treatment (ChiSq 6 = 3.0146, p=0.807)

# Graph predictions based on m2 to estimate effect size

weight1 <- emmip(m.weight1, ~trt_label, type = "response", CIs = TRUE, plotit = FALSE)
weight2 <- emmip(m.weight2, ~trt_label, type = "response", CIs = TRUE, plotit = FALSE)

emplot_weight <- emmip(m.weight2, ~trt_label, type = "response", CIs = TRUE, plotit = FALSE)
emplot_weight <- emplot_weight %>% 
  rename(
    post_kgs_C = yvar
  )


emplot_weight <- emplot_weight %>% 
  mutate(
    trt_label = fct_relevel(trt_label, "Control", "Apivar", "Amitraz EC", "OA Dribble", "5x OA Dribble", "OA Vapor", "HopGuard")
  )


```


## Plot post-harvest colony weight 

```{r}
paneB <- ggplot() +
  geom_boxplot(aes(x = trt_label, y = post_kgs_C), fill = "#031C38", outlier.shape = NA, color = "black", data = temp2) +
  geom_point(aes(x = trt_label, y = post_kgs_C), shape=20, size=10, color="black", fill="black", data = emplot_weight) +
  #xlab("Treatment") +
  ylab("\nHive weight (kg)") +
  theme_classic( base_size = 30) +
  theme(axis.text.x=element_text(angle=45,hjust=1), axis.title.x=element_blank())

# Color

paneB2 <- ggplot() +
  geom_boxplot(aes(x = trt_label, y = post_kgs_C, fill = trt_type), outlier.shape = NA, color = "black", data = temp2) +
  geom_point(aes(x = trt_label, y = post_kgs_C), shape=20, size=10, color="black", fill="black", data = emplot_weight) +
    scale_colour_manual(
    limits = c("Control", "Synthetic", "Natural"),
    labels = c("Control", "Synthetic", "Natural"),
    values = c("white", "#F8766D", "#619CFF"), 
    aesthetics = "fill"
  ) +  
  #xlab("Treatment") +
  ylab("\nHive weight (kg)") +
  theme_classic( base_size = 30) +
  theme(axis.text.x=element_text(angle=45,hjust=1), axis.title.x=element_blank(), legend.position = "none")

```



## Combine plot panes

```{r}
# version 1
ggarrange(paneA, paneB, ncol = 1, nrow = 2, labels = "AUTO", font.label = list(size = 30))

ggsave("outputs/strength_weight v1.png", width = 8, height = 17.25, units = "in")

# version 2
ggarrange(paneA2, paneB2, ncol = 1, nrow = 2, heights = c(0.82, 1), labels = "AUTO", font.label = list(size = 30)) # common.legend = TRUE

ggsave("outputs/strength_weight v2.png", width = 8, height = 17.25, units = "in")

# version 3 - preferred version
ggarrange(paneA3, paneB2, ncol = 2, nrow = 1, labels = "AUTO", font.label = list(size = 30)) # common.legend = TRUE

ggsave("outputs/strength_weight 2024-04-24.png", width = 17.25, height = 8, units = "in")
```


